## Setup

```{r}
source('../macrophage_figure_setting.R')

plot_dir <- '/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/04_results/'

```

# Whole proteome visualization

## PCA

```{r, fig.width=2, fig.height=2}
# pca_plot() and pc_label() are centralized in ../macrophage_figure_setting.R;
# axis %s are read from percent_explained.csv (written by src/pca_utils.py)
pca_dir <- paste0(plot_dir, "pca_plots/")

read.csv(paste0(pca_dir, "pca_results.csv")) -> df
read.csv(paste0(pca_dir, "loadings_df.csv")) -> loadings_df

pca_plot(
  df,
  loadings_df,
  "principal.component.1",
  "principal.component.2",
  pc_label(pca_dir, "PC1"),
  pc_label(pca_dir, "PC2"),
  paste0(pca_dir, "pca"),
  arrow_scaling = 150
)





```

## RNA PCA
```{r, fig.width=2, fig.height=2}
pca_dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/02_RNA-seq/PCA_5000_genes/'

read.csv(paste0(pca_dir, "principal_components.csv")) -> df
read.csv(paste0(pca_dir, "loadings_df.csv")) -> loadings_df

pca_plot(
  df, 
  loadings_df,
  "PC1",
  "PC2",
  pc_label(pca_dir, "PC1"),
  pc_label(pca_dir, "PC2"),
  paste0(pca_dir, "pca"),
  arrow_scaling = 1000
)
  
```

## GSEA of PC loadings

```{r, fig.height=3, fig.width=3.2}
clean_term <- function(term) {
  str_split_fixed(term, "_", 2)[2] -> term
  gsub("\\b([a-z])", "\\U\\1", term, perl=TRUE) -> term
  str_replace_all(term, "_", " ") -> term
  str_wrap(term, 40) -> term
  return(term)
}
for (pc in c("PC1", "PC2")) {
  paste0(plot_dir,
         "pca_plots/",
         pc,
         "/gseapy.gene_set.prerank.report.csv") %>%
    read_csv(show_col_types = FALSE) %>%
    mutate(
      Term = lapply(Term, FUN = clean_term) %>% unlist(),
      `-log10(FDR)` = -log10(`FDR q-val`)
    ) %>%
    arrange(NES) -> df
  
  max_score  <- max(df$`-log10(FDR)`[is.finite(df$`-log10(FDR)`)], na.rm = TRUE)
  df$`-log10(FDR)`[is.infinite(df$`-log10(FDR)`)] <- 5.5
  
  
  df <- df %>%
    filter(`FDR q-val` < 0.01) 
  rbind(
    df %>% head(8),
    df %>% tail(8)
  ) %>% distinct() %>%
    tail(15) %>%
    ggplot(aes(
      # x = NES (not -NES) so +NES sits on the +PC side, matching the PCA panel orientation
      x = NES,
      y = reorder(Term, NES),
      fill = `-log10(FDR)`
    )) +
    geom_col(color = "black",
             size = LINE_WIDTH,
             width = 0.5) +
    my_theme() +
    geom_vline(xintercept = 0,
               color = "black",
               size = LINE_WIDTH) +
    labs(x = paste0(pc, " NES"),
         y = NULL,
         fill = expression("-log"[10] * "(FDR)")) +
    theme(
      axis.text.y = element_text(size = 5, family = FONT_FAMILY),
      legend.text = element_text(margin = margin(l = 1)),
      legend.key.height = unit(4, "mm"),
      legend.key.width = unit(4, "mm"),
      panel.grid.major = element_line(color = "grey90", size = LINE_WIDTH),
    ) + 
    scale_fill_continuous(low = "#B3CDE3", high = "#B10F7C", limits = c(1.5, 5.5)) -> plt
  print(plt)
  save_plot(
    paste0(plot_dir, "pca_plots/", pc, "/top_enriched_sets"),
    width = 3.2,
    height = 2.2
  )
} 

plot_dir
```



## Volcano plots

```{r}
volcano_dir <- paste0(plot_dir, "volcano_plots/")
paste0(volcano_dir, "volcano_data_long_format.csv") %>% read.csv() -> volcano_data
volcano_dir
```

```{r, fig.width=2, fig.height=2.2}

other_str <- "Other"
nudge_x = 1
nudge_y = 0
label_bound = 0.7
label_upper_bound = NA

y_max <- max(abs(volcano_data$X.log10_pval)) * 1.2
x_max <- max(abs(volcano_data$log2_FC)) * 1.1



make_condition_plot <- function(
    category, 
    category_proteins,
    condition, 
    volcano_data, 
    control_condition, 
    highlight_proteins = c(), 
    highlight_color = "#21918c"){
  x_var <- "log2_FC"
  y_var <- "X.log10_pval"
  change_var = "Regulation"
  name_var = "protein"
  group <- category
  file_name <- category_proteins
  title <- paste0(control_condition," vs ", condition)
  
  cols_binary <- c("Other" = other_color)
  cols_binary[[group]] <- highlight_color
  alphas_binary <- c("Other" = 0.5)
  alphas_binary[[group]] <- 1
  alphas_binary <- c(0.2,0.9)
  
  # Important to label all protein lists with a "uniprot" column 
  read.csv(file_name) %>% 
    select(uniprot) %>% 
    unique() -> 
    functional_proteins
  functional_proteins <- 
    intersect(functional_proteins$uniprot, volcano_data$uniprot)
  
  # Label all proteins belonging to our functional category
  volcano_data %>% 
    mutate(group = is.element(uniprot, functional_proteins)) -> 
    volcano_data
  
  sig <- volcano_data[ volcano_data[[change_var]] %in% c("Significant Up", "Significant Down") , ]
  functional_and_sig <- sig[sig[["group"]] == TRUE, ]
  label_data <- sig[ sig[[name_var]] %in% highlight_proteins , ]
  
  # quadrant count labels
  functional_and_sig[functional_and_sig[[change_var]] == "Significant Up",] %>% nrow() -> func_sig_up
  functional_and_sig[functional_and_sig[[change_var]] == "Significant Down",] %>% nrow() -> func_sig_down
  nonfunctional_and_sig <- volcano_data[volcano_data[["group"]] == FALSE, ]
  nonfunctional_and_sig[nonfunctional_and_sig[[change_var]] == "Significant Up",] %>% nrow() -> nonfunc_sig_up
  nonfunctional_and_sig[nonfunctional_and_sig[[change_var]] == "Significant Down",] %>% nrow() -> nonfunc_sig_down
  
  # fractions to scale placement of labels with limits of plot
  label_x <- x_max * (4.7/5.3)
  x <- c(-label_x,label_x)
  y <- rep(
    y_max * (15.5/16),2)
  label <- c(func_sig_down, func_sig_up)
  text_labels <- data.frame(x, y, label)
  
  y <- y - (y_max * (1.5/16))
  label <- c(nonfunc_sig_down, nonfunc_sig_up)
  overall_labels <- data.frame(x, y, label)
  
  
  # Plot data
  volcano_data %>%
    # label all proteins that are in our functional category and are significant
    mutate("f_and_sig" = uniprot %in% 
             functional_and_sig$uniprot) %>%
    mutate(significant = uniprot %in% sig$uniprot) %>%
    arrange(f_and_sig) -> volcano_data
  volcano_data$group <- apply(X=volcano_data, MARGIN = 1, FUN = is_in_group, fun_group=group)
  volcano_data$group <- factor(volcano_data$group, levels = c("Other", group))
  volcano_data %>%
    arrange(group) %>%
    ggplot(aes_string(x=x_var,
               y=y_var,
                fill = "group",
               alpha = "significant")) +
    ylim(0,y_max) +
    xlim(-x_max,x_max)+
    scale_alpha_manual(values = alphas_binary, guide="none") +
    geom_point(data = volcano_data %>% filter(significant == FALSE),
           size = 1,
           shape=21,
           stroke=POINT_STROKE / 2,
           col="black",
           fill = 'gray88'
           ) +
    geom_point(data = volcano_data %>% filter(group == "Other", significant == TRUE), 
               size = 1,
               shape=21,
               stroke=POINT_STROKE / 2,
               col="black") +
    geom_point(data = volcano_data %>% filter(group != "Other", significant == TRUE),
      size = 1,
               shape=21,
               stroke=POINT_STROKE / 2,
               col="black") +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", size=LINE_WIDTH) +
    geom_vline(xintercept = c(-FC_CUTOFF, FC_CUTOFF), linetype = "dashed", size=LINE_WIDTH) +
    geom_text_repel(
      data = label_data[label_data[[change_var]] == "Significant Down",],
      aes_string(x = x_var,
      y = y_var,
      # label with gene name instead of uniprot
      label = name_var),
      direction    = "both",
      color = "black",
      box.padding = 1,
      nudge_x = -nudge_x,
      nudge_y = nudge_y,
      xlim = c(-label_upper_bound, -label_bound),
      size =FONT_SIZE_MM,
      #bg.color = "white",
      max.overlaps = getOption("ggrepel.max.overlaps", default = 30),
      min.segment.length = 0,
      segment.size = LINE_WIDTH,
      inherit.aes = FALSE) +
    geom_text_repel(
      data = label_data[label_data[[change_var]] == "Significant Up",],
      aes_string(x = x_var,
      y = y_var,
      # label with gene name instead of uniprot
      label = name_var),
      direction    = "both",
      color = "black",
      box.padding = 1,
      nudge_x = nudge_x,
      nudge_y = nudge_y,
      xlim = c(label_bound,label_upper_bound),
      force_pull = 0,
      max.overlaps = getOption("ggrepel.max.overlaps", default = 30),
      size = FONT_SIZE_MM,
      #bg.color = "white",
      min.segment.length = 0,
      segment.size = LINE_WIDTH,
      inherit.aes = FALSE) +
    theme_test() +
    labs(
      x = substitute(paste('protein expression, log'[2], "(", n, '/', m, ")"), list(n = condition, m = control_condition)),
      y = substitute(paste('-log'[10], '(p-value)'))
    ) +
    scale_fill_manual(values=cols_binary, breaks = c(group, "Other")) +
    my_theme() +
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.title=element_blank(),
      legend.box.background = element_blank(),
      legend.position = "top",
      aspect.ratio=1,
      legend.key = element_blank(),
      legend.margin = margin(l = -5, b = -5.5),
      legend.box.margin = margin(l = -5, b = -5.7),
      legend.key.spacing.y = unit(-1, "cm"),
      legend.key.spacing.x = unit(-0.05, "cm")) + 
  geom_label(inherit.aes=FALSE,
          data=text_labels,
          mapping=aes(x=x,y=y,label=label),
          size=6,
          size.unit="pt",
          show.legend = FALSE,
          color='white', fill=highlight_color) +
  geom_label(inherit.aes=FALSE,
      data=overall_labels,
      mapping=aes(x=x,y=y,label=label),
      size=6,
      size.unit="pt",
      show.legend = FALSE,
      color='white', fill="lightgrey") -> v_plot
  print(v_plot)
  
  # make directory
  dir.create(
    paste0(volcano_dir, "functional_groups_highlighted/", group),
    showWarnings = FALSE
  )
  
  save_plot(paste0(volcano_dir, "functional_groups_highlighted/", group, "/", title), 
            width=2, 
            height=4)
}

```









## Multiple condition volcano plot

```{r, fig.width=2, fig.height=4}
is_in_group <- function(row, fun_group){
  if (row["group"] == fun_group) {
    return(fun_group)
  } else {
    return("Other")
  }
}
options(ggrepel.max.overlaps = Inf) # for crowded volcano plots


cols <- c(
  "M0" = "#AAAAAA",
  "TLR1-2"= "#EC2427",
  "TLR3" = "#F58420",
  "TLR4" = "#FFCC31",
  "TLR7" = "#4869B2",
  "TLR8" = "#5CBED7",
  "TLR9" = "#237D41",
  "STING" = "#742D16"
)

color_code <- c(
  "Mitochondrial translation" ="#237D41",
  "Mitochondrial" = "#5CBED7",
  "Other" = 'gray88'
)


group_hierarchy <- function(df){
  df %>%
    mutate(group = 
           case_when(
             `Mitochondrial translation` == TRUE ~ "Mitochondrial translation",
             `Mitochondrial` == TRUE ~ "Mitochondrial",
             TRUE ~ "Other"
           )) %>% 
    return()
}

volcano_df_fn <- '/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/04_results/volcano_plots/volcano_data_long_format.csv'
volcano_df_fn |> 
  read_csv(show_col_types = FALSE) |> 
  group_hierarchy() -> volcano_data

for (cond in conditions){
 make_condition_plot_multi(
  condition = cond,
  volcano_data = volcano_data %>% filter(condition == cond),
  control_condition = "M0",
  color_code = color_code,
  highlight_proteins = c("FASTKD2","MTG1","PTCD1","DAP3","QRSL1","GATC","ACOD1","RSAD2","NOCT","IFIH1","SOD2","AK4","TRAP1")
)
}
group_hierarchy <- function(df){
  df %>%
    mutate(group = 
           case_when(
             `Electron transport chain` == TRUE ~ "ETC",
             `Glycolysis` == TRUE ~ "Glycolysis",
             TRUE ~ "Other"
           )) %>% 
    return()
}
'/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/04_results/volcano_plots/volcano_data_long_format.csv' %>% 
  read_csv(show_col_types = FALSE) %>%
  group_hierarchy() -> volcano_data

color_code <- c(
  "ETC" = "#4869B2",
  "Glycolysis" = "#742D16",
   "Other" = 'gray88'
)
for (cond in conditions){
 make_condition_plot_multi(
  condition = cond,
  volcano_data = volcano_data %>% filter(condition == cond),
  control_condition = "M0",
  color_code = color_code,
  highlight_proteins = c("HK2","PFKP","GAPDH","PGK1","PGM1","ENO1","PKM","LIPA","NDUFA4","NDUFS1","MT-ND2","SDHA","CYCS","MT-CO1","COX5A","IDO1", "GPI", "ALDOA", "TPI1")
)
}


```


```{r}
volcano_data |> 
  filter(protein %in% c("FASTKD2","MTG1","PTCD1","DAP3","QRSL1","GATC","ACOD1","RSAD2","NOCT","IFIH1","SOD2","AK4","TRAP1"), condition == "TLR4" ) 

c("FASTKD2","MTG1","PTCD1","DAP3","QRSL1","GATC","ACOD1","RSAD2","NOCT","IFIH1","SOD2","AK4","TRAP1") %in% volcano_data$protein
```


## Upset plot
```{r, fig.height=2.2, fig.width=2.75}
library(patchwork)
library(ComplexUpset)
library(ggplot2)

volcano_dir <- paste0(plot_dir, "volcano_plots/")
paste0(volcano_dir, "volcano_data_long_format.csv") %>% read.csv() -> volcano_data


# build sets iteratively
sets <- list()
for(cond in conditions){
  if (cond == "TLR9"){
    next
  }
 for (direction in c("Up", "Down")){
   sets[[paste(cond, direction)]] <- volcano_data %>% filter(
     condition == cond,
     Regulation == paste0("Significant ", direction)
   ) %>% pull(uniprot)
 }
}

# transform to binary data frame
all_elements <- unique(unlist(sets))
df <- as.data.frame(sapply(sets, function(set) {
  all_elements %in% set
}))

# define plot aesthetics
upset_plot_theme <-
  theme_classic() +
  theme(
    axis.text.y = TEXT_ELEMENT,
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_blank(),
    axis.title.y = TEXT_ELEMENT,
    panel.grid = element_blank(),
    axis.line.x = AXIS_LINE,
    axis.line.y = AXIS_LINE
  )

upset(
  df,
  sort_sets = FALSE,
  intersect = rev(names(sets)),
  set_sizes = FALSE,
  n_intersections = 10,
  matrix = (intersection_matrix(
    geom = geom_point(
      size = 1
    ),
    segment = geom_segment(
      linewidth = LINE_WIDTH
    )
  )),
  annotations = list(
    "intersection size" = intersection_size(
      counts = TRUE,
      color = "black",
      size = LINE_WIDTH,
      show.legend = FALSE,
      width = 0.65,
      fill = cols["TLR8"],
      text = list(size = FONT_SIZE_MM, color = "black", family = FONT_FAMILY, vjust = -0.1, check_overlap = TRUE)
    ) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.2)))
  ),
  base_annotations = list(),
  stripes = "white",
  min_size = 10,
  themes = list(
    `intersection size` = upset_plot_theme + theme(axis.ticks.y = AXIS_LINE),
    intersections_matrix = upset_plot_theme + theme(axis.line.x = element_blank(), axis.line.y = element_blank(), axis.title.y = element_blank())
  )
) + patchwork::plot_layout(heights = c(0.5, 0.7))
ggsave(paste0(plot_dir, "upset_plot.svg"), height = 2.2, width = 2.75)
```







## heatmap
```{r}
paste0(plot_dir, "../","03_combined_files/1/04_combcond_forpca_channelratio_or_rawsignal_wp_keratins_removed_min4rep_filter.csv") %>% 
  read_csv() -> wp_data

wp_data$protein <- make.unique(wp_data$protein)

wp_data %>%
  select(-c("description","uniprot","...1", "M0")) %>%
  column_to_rownames("protein") -> wp_data
wp_data %>%
  apply(1,scale) %>% 
  t() %>% 
  as.data.frame()  %>% 
  as.matrix() -> wp_scaled_data
colnames(wp_scaled_data) <- wp_data %>% colnames()
ht_data <- wp_scaled_data

wp_scaled_data <- ht_data
  n_markers <- nrow(wp_scaled_data)
    wp_scaled_data %>% 
    colnames() -> cls 

  
  wp_scaled_data <- wp_scaled_data[,
                                   order(factor(colnames(wp_scaled_data), levels = names(cols)))
                                   ]
  ba <- HeatmapAnnotation(conditions=colnames(wp_scaled_data), 
                          col=list(conditions=cols[names(cols) != "M0"]),
                          show_legend = FALSE,
                          show_annotation_name = FALSE)
  

  panel_fun = function(index, nm) {
    grid.rect()
    grid.text(paste0(length(index), " rows"), 0.5, 0.5)
  }
wp_scaled_data %>%  as.data.frame() %>% rownames_to_column(var = "protein") %>% write_csv(paste0(plot_dir, "heatmap/wp_exhaustion_markers_median_combined_scaled_data.csv"))
```


```{r, fig.width=3, fig.height=6}

  wp_scaled_data %>% 
    Heatmap(name="zscore", 
            cluster_rows = TRUE, 
            bottom_annotation = ba,
            cluster_columns = FALSE, 
            show_row_names = FALSE,
            row_names_side = "left",
            show_column_names = TRUE,
            #rect_gp = gpar(col = "white", lwd = LINE_WIDTH),
            show_column_dend = FALSE,
            show_row_dend = TRUE,
            row_km = 6,
            row_names_gp = gpar(fontsize =6 ),
            row_title_gp = gpar(fontsize =6 ),
            row_title_side = "right",
            row_gap = unit(2, "mm"),
            col = colorRamp2(c(-2, 0, 2), c("blue", "white", "red")),
            show_heatmap_legend = TRUE,
            heatmap_legend_param = list(
              legend_height = unit(1, "cm"), 
              grid_width = unit(0.2, "cm"), 
              title_gp =gpar(fontsize =6 ), 
              labels_gp = gpar(fontsize =6 ) )
            ) -> ht

pdf(file=paste0(plot_dir, "heatmap/median_combined_scaled.pdf"),width=2,height=2.3)
print(ht)
dev.off()
print(ht)
```




## WP vs RNA-Seq correlation plots

```{r, fig.height=2, fig.width=3}

correlation_dir <- paste0(plot_dir, "wp_vs_rnaseq/")

correlation_cols <-
  c(
    "light_grey" = "lightgrey",
    "dark_grey" = "grey40",
    "teal" = "#21918c",
    "violet" = "#440154",
    "yellow" = "#bade28",
    'black' = 'black'
  )

create_wp_rna_corr_plot <-
  function(comparison,
           proteins_to_label,
           x_lim = 4.6,
           y_lim = 10) {
    condition = strsplit(comparison, "_")[[1]][1]
    control = strsplit(comparison, "_")[[1]][2]
    read.csv(paste0(correlation_dir, condition, " vs. ", control, ".csv")) %>%
      mutate(gene_color = expression)  |> select(protein, log2FoldChange, logFC_wp, gene_color) |> distinct()-> df
    
    y_lim = max(abs(df$log2FoldChange)) * 1.2
    y_min = min(df$log2FoldChange) * 1.2
    x_lim = max(abs(df$logFC_wp)) * 1.2
    
    df %>% filter(protein %in% proteins_to_label) -> label_data
    
    label <- df %>%
      summarise(
        left_red = sum(gene_color == 'yellow' & logFC_wp < 0),
        right_red = sum(gene_color == 'yellow' & logFC_wp > 0),
        left_cyan = sum(gene_color == 'teal' & logFC_wp < 0),
        right_cyan = sum(gene_color == 'teal' & logFC_wp > 0),
        left_yellow = sum(gene_color == 'dark_grey' & logFC_wp < 0),
        right_yellow = sum(gene_color == 'dark_grey' & logFC_wp > 0),
        down_blue = sum(gene_color == 'violet' & log2FoldChange < 0),
        up_blue = sum(gene_color == 'violet' & log2FoldChange > 0),
        grey = sum(gene_color == 'light_grey')
      ) %>% unlist()
    
    x_lbl_sprd = x_lim * 0.915
    y_lbl_sprd = y_lim * 0.98
    y_lbl_min = y_min * 0.98
    lbl_alpha = 1
    
    
    x <-
      c(-x_lbl_sprd,
        x_lbl_sprd,
        -x_lbl_sprd,
        x_lbl_sprd,
        -x_lbl_sprd,
        x_lbl_sprd,
        0,
        0,
        0)
    y <-
      c(0,
        0,
        y_lbl_sprd,
        y_lbl_min,
        y_lbl_min,
        y_lbl_sprd,
        y_lbl_min,
        y_lbl_sprd,
        0)
    alpha <- c(rep(lbl_alpha, 8), 0)
    color <-
      c(
        'yellow',
        'yellow',
        'teal',
        'teal',
        'dark_grey',
        'dark_grey',
        'violet',
        'violet',
        'black'
      )
    text_labels <- data.frame(x, y, label, color)
    ratio <- paste0("(", condition, "/", control, ")")
    pearson_r_color <- 'grey40'
    df %>%
      ggplot(aes(x = logFC_wp, y = log2FoldChange)) +
      geom_point(
        alpha = 0.2,
        size = 1,
        mapping = aes(fill = gene_color),
        shape = 21,
        stroke = 0
      ) +
      geom_smooth(
        method = "lm",
        se = FALSE,
        color = pearson_r_color,
        formula = 'y ~ x',
        size = LINE_WIDTH * 4
      ) +
      labs(x = substitute(paste('protein expression, log'[2], n), list(n = ratio)),
           y = substitute(paste('RNA expression, log'[2], n), list(n = ratio))) +
      theme_bw() +
      geom_hline(
        yintercept = c(FC_CUTOFF,-FC_CUTOFF),
        linetype = "dashed",
        size = LINE_WIDTH
      ) +
      geom_vline(
        xintercept = c(FC_CUTOFF,-FC_CUTOFF),
        linetype = "dashed",
        size = LINE_WIDTH
      ) +
      theme(aspect.ratio = 1, legend.position = "none") +
      
      # specific proteins
      geom_point(
        mapping = aes(fill = gene_color),
        alpha = 1,
        data = label_data,
        size = 1,
        shape = 21,
        color = "black",
        stroke = POINT_STROKE / 2
      ) +
      geom_text_repel(
        data = label_data,
        mapping = aes(label = protein, color = gene_color),
        size = FONT_SIZE_MM,
        segment.color = "black",
        segment.size = LINE_WIDTH,
        min.segment.length = 0.2,
        #bg.color = "white",
        box.padding = 1,
        direction = "both"
      ) +
      scale_color_manual(values = correlation_cols) +
      my_theme() +
      theme(aspect.ratio = 1,
            legend.position = "none") +
      geom_label(
        inherit.aes = FALSE,
        data = text_labels,
        mapping = aes(
          x = x,
          y = y,
          label = label,
          fill = color,
          alpha = alpha
        ),
        color = 'white',
        size = 6,
        label.size = 0,
        size.unit = "pt"
      ) +
      scale_fill_manual(values = correlation_cols) +
      ylim(y_min, y_lim) +
      xlim(-x_lim, x_lim) +
      stat_cor(
        aes(label = ..r.label..),
        color = pearson_r_color,
        label.x = -(x_lim * 0.95),
        label.y = (y_lim * 0.67),
        size = FONT_SIZE_MM
      ) -> plot
    print(plot)
    
    save_plot(
      paste0(
        correlation_dir,
        condition,
        "_",
        control,
        "_pearson_corrplot"
      ),
      height = 2,
      width = 2
    )
  }

paste0(volcano_dir, "volcano_data_long_format.csv") %>%
  read_csv(show_col_types = FALSE) -> volcano_data
for (macrophage_condition in conditions) {
  create_wp_rna_corr_plot(
    comparison = paste0(macrophage_condition, "_", control_condition),
    proteins_to_label = c("CAMK1","CLIC1","S100A4","SCO2","ENO1","PGM2","PFKL","PKM","CANX","DNAJA2","HSP900A1","HSPA9","PFDN2","EMC1","UTP11","UTP14A","SERBP1","FABP3","GATD1","ADAM17","VTN","ERAL1","SDF4","PYCR2", "STAT4")
  )
}


correlation_dir
```



```{r, fig.width=1.5, fig.height=1.5}

column_gap <- 0.5/2
text_graphics <- gpar(fontsize=6, fontfamily = "Arial")
c(
  "TLR1-2" = 0.4,
  "TLR3" = 0.42,
  "TLR4" = 0.42,
  "TLR7" = 0.31,
  "TLR8" = 0.41,
  "TLR9" = 0.04,
  "STING" = 0.45
) %>% 
  as.matrix() -> heatmap_data
heatmap_data %>%
  Heatmap(
          cluster_rows = FALSE, 
          cluster_columns = FALSE, 
          show_row_names = TRUE,
          show_column_names = FALSE,
          rect_gp = gpar(col = "white", lwd = LINE_WIDTH* 4),
          show_column_dend = FALSE,
          show_row_dend = FALSE,
          column_title = "RNA vs protein correlation",
          heatmap_legend_param = list(
            title = "Correlation\ncoefficient (R)",
            legend_height = unit(1.75,"cm"),
            title_gp = text_graphics,
            labels_gp = text_graphics,
            border = "transparent",
            grid_width = unit(1, "mm")
          ),
          row_names_side = "left",
          column_gap = unit(column_gap, "mm"),
          column_title_gp = text_graphics,
          row_names_gp = gpar(
            fontsize = 6,
            fontfamily = "Arial"
          ),
          cell_fun =function(j, i, x, y, width, height, fill) {
        grid.text(sprintf("%.2f", heatmap_data[i, j]), x, y, gp = gpar(fontsize = 6))
},
    col=colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))
  ) -> ht

pdf(
  paste0(correlation_dir, "all_conditions_correlation.pdf"),
  width = 1.5,
  height = 1.5
)
print(ht)
dev.off()
print(ht)
```

```{r}
lfc_cutoff = log2(1.5)
p_value_cutoff = 0.05
df <- data.frame()
rna_de_path = '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/RNA-seq/RNA-seq_Beatrice/macrophage_analysis/DEG/'
conditions <- names(c(
    "TLR1-2"= "#EC2427",
  "TLR3" = "#F58420",
  "TLR4" = "#FFCC31",
  "TLR7" = "#4869B2",
  "TLR8" = "#5CBED7",
  "TLR9" = "#237D41",
  "STING" = "#742D16"
))
for(macrophage_condition in conditions){
  
  # RNA results
  read_excel(path = paste0(rna_de_path, "M0lowvs", str_split(macrophage_condition, "-")[[1]][1], "/DE.xlsx"),
             sheet = "DEseq2, coding") %>%
    filter(pvalue < p_value_cutoff) -> de_results
  de_results %>% filter(log2FoldChange > lfc_cutoff) %>% nrow() -> n_up
  de_results %>% filter(log2FoldChange < -lfc_cutoff) %>% nrow() -> n_down
  df <- rbind(df, 
                list(
                  "condition" = macrophage_condition,
                  "n_changes" = n_up,
                  "direction" = "upregulated",
                  "dataset" = "RNA"
                  ))
  df <- rbind(df, 
                list(
                  "condition" = macrophage_condition,
                  "n_changes" = n_down,
                  "direction" = "downregulated",
                  "dataset" = "RNA"
                  )
              )
    
    # protein results
    volcano_data %>% filter(p_value < p_value_cutoff, condition == macrophage_condition) -> protein_de_results
    protein_de_results %>% filter(log2_FC > lfc_cutoff) %>% nrow() -> n_up
    protein_de_results %>% filter(log2_FC < -lfc_cutoff) %>% nrow() -> n_down
    
      df <- rbind(df, 
                list(
                  "condition" = macrophage_condition,
                  "n_changes" = -n_up,
                  "direction" = "upregulated",
                  "dataset" = "Protein"
                  ))
      df <- rbind(df, 
                list(
                  "condition" = macrophage_condition,
                  "n_changes" = -n_down,
                  "direction" = "downregulated",
                  "dataset" = "Protein"
                  )
        )
}
df$dataset <- factor(df$dataset, levels = c("Protein", "RNA"))
```




## bar plots

```{r, fig.height=2.5, fig.width= 2.5}
x_axis_exand <- 0.29
dodge_fun <- position_dodge(width = 0.8)
df %>%
  ggplot(aes(x = n_changes, y = factor(condition, levels = rev(conditions)), fill=direction, label = abs(n_changes))) + 
  geom_bar(stat = "identity", position = dodge_fun, width = 0.5) + 
  geom_text(
    data = df %>% filter(dataset == "Protein"),
    position = dodge_fun, size = FONT_SIZE, size.unit = "pt", hjust = 1.1) +
    geom_text(
    data = df %>% filter(dataset == "RNA"),
    position = dodge_fun, size = FONT_SIZE, size.unit = "pt", hjust = -0.1) +
  facet_wrap(~dataset, scales="free_x") + 
  my_theme() + 
  scale_fill_manual(values = c("up (FC > 1.5)" = SIG_UP_COL, "down (FC < -1.5)" = SIG_DOWN_COL)) + 
  theme(
    panel.spacing = unit(0, "mm"),
    legend.position = "bottom",
    legend.margin = margin(t = -8, b = 0),
    legend.key.size = unit(3, "mm"),
    legend.key.spacing = unit(3, "mm"),
              legend.text = element_text(
            size=6, color = "black", 
            family="Arial",
            margin = margin(l =0)),
    strip.background = element_blank(),
    strip.text = element_text(
    size=FONT_SIZE, 
    color = "black", 
    family=FONT_FAMILY))+ 
  scale_x_continuous(expand = c(0,0)) +
  labs(x = "expression changes (p-value < 0.05)",y = NULL, fill = NULL ) +
  facetted_pos_scales(
    x = list(
      scale_x_continuous(expand = expansion(mult = c(x_axis_exand,0), 
                                            ), labels = function(x) signif(abs(x), 3)),
      scale_x_continuous(expand = expansion(mult = c(0,x_axis_exand)))
      )
    )
ggsave(
  paste0(correlation_dir, "all_conditions_n_changes.svg"),
  width = 2.5,
  height = 2.5
)
```


```{r, fig.width=2, fig.height=2}
expression_bar_graph <- function(
    proteins, 
    conditions,
    control_condition,
    width,
    height){
paste0('/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/01_unenriched proteomics/03_results/20250219_3reps/03_combined_files/1/02_combfiles_percctrl_wp.csv') %>%
    read_csv(show_col_types = FALSE) %>% 
    filter(protein %in% proteins) %>% 
    column_to_rownames("protein") -> df
  df[,colnames(df)[grepl(
    paste(conditions, collapse = "|"), colnames(df)
    )]] %>% 
    t() %>% as.data.frame() %>%
    rownames_to_column("replicate") %>%
    rowwise() %>%
    mutate(condition = str_split(replicate, "_")[[1]][1]) %>%
    pivot_longer(
      cols = proteins,
      names_to = "protein",
      values_to = "percent_control"
    ) %>%
    drop_na() -> df
  
  print(df)
  # Calculates mean, sd, se and IC
  df %>%
    group_by(condition, protein) %>%
    summarise( 
      n=n(),
      mean=mean(percent_control),
      sd=sd(percent_control),
      fun = mean
    ) %>%
    mutate(se=sd/sqrt(n))  %>%
    mutate(ic=se * qt((1-0.05)/2 + .5, n-1)) -> stat_df
  df$condition <- factor(df$condition, levels = conditions)
  df %>% 
    ggplot(
    aes(x =condition, y = percent_control, fill = condition)
  ) + 
    geom_bar(stat = "summary", 
           fun = "mean",
           show.legend = FALSE,
           width = 0.7,
           alpha = 0.7) +
    ggbeeswarm::geom_beeswarm(shape = 21, 
              color = "black",
              size = POINT_SIZE - 1,
              stroke = POINT_STROKE,
              show.legend = FALSE) + 
    geom_errorbar(data = stat_df, 
                mapping = aes(
                  x = condition, 
                  ymin = mean - se,
                  ymax = mean + se
                ),
                inherit.aes = FALSE,
                width = 0.4,
                size = LINE_WIDTH) +
    facet_wrap(~protein, scales = "free_y") + 
    my_theme() + 
    geom_hline(yintercept = 100, 
               size = LINE_WIDTH, 
               linetype="dashed"
               ) + 
  scale_y_continuous(
    #limits = c(0, 180),
    expand = expansion(mult = c(0, 0.1))
  )  + 
  theme(
    strip.background = element_blank(),
    panel.spacing = unit(2,"mm"),
    strip.text = element_text(size = 8),
    panel.border =element_blank(),
    axis.line.x = AXIS_LINE,
    axis.line.y = AXIS_LINE
  ) + 
      theme(axis.text.x = element_text(
        color="black",angle=45, size=FONT_SIZE, family="Arial", vjust = 1, hjust=1)) +
  scale_fill_manual(values = cols) +
  labs(x = "", y = paste0("protein expression (%",control_condition,")")) -> plt
  print(plt)
  
  save_plot(    
    paste0(plot_dir, 
           "bargraphs/",
           paste(proteins, collapse = "_"),
           "_expression_bargraph"),
    width = width,
    height = height)
}
plot_dir
expression_bar_graph(proteins = c("DDX3X"), conditions = names(cols), control_condition = "M0", width = 2, height = 2)
```


# INFG heatmap





```{r, fig.width=2.7, fig.height=2.3}
ifng_proteins <- msigdb_genes("HALLMARK_INTERFERON_GAMMA_RESPONSE", "H")
protein_list <- ifng_proteins

wp_with_heatmap_labels <- function(protein_list, proteins_to_mark = NULL, results_name, mark_top_n = NULL){
'/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/03_combined_files/1/06_combfiles_forpca_channelratio_or_rawsignal_wp_keratins_removed_min4rep_filter.csv' %>%
  read_csv(show_col_types = FALSE) -> df
label_pattern = "pepNum"
df %>% colnames() -> l
df %>% 
  # drop several columns we will not use for the heatmap
  select(l[!grepl(label_pattern, l)]) %>%
  select(-c("...1", "description", "uniprot")) %>%
  mutate(protein = make.unique(protein)) %>%
  column_to_rownames("protein") %>% 
  drop_na() -> df
df %>% colnames() -> l
lfc_df<- log2(df / rowMeans(df[l[grepl("M0", l)]]))

lfc_df <- lfc_df[l[!grepl("M0", l)]]
lfc_df %>% colnames() -> l
lfc_df <- lfc_df[l[!grepl("TLR9", l)]]

colnames(lfc_df) %>% 
  lapply(FUN=function(name){str_split(name, "_")[[1]][1]}) %>% 
  unlist() -> channel_conditions
channel_conditions <- factor(channel_conditions, levels= names(cols))


colnames(lfc_df) <- colnames(lfc_df) %>%
  lapply(FUN=function(name){str_split(name, "_processed")[[1]][1]}) %>% 
  unlist() 

txt_gp <- gpar(col = "white", fontsize = 6, font="Arial")

ba <- HeatmapAnnotation(
  empty = anno_empty(height = unit(3, "mm"),border = FALSE),
  conditions= anno_block(
        gp = gpar(fill = cols[!grepl("M0|TLR9", names(cols))], col = "transparent"),  # Set border color same as fill
        labels = names(cols)[!grepl("M0|TLR9", names(cols))],
        height=unit(3,"mm"),
        which = "column",
        labels_gp = txt_gp)
  )
ra <- HeatmapAnnotation(
  cluster = anno_block(
    labels = c("Cluster 1", "Cluster 2", "Cluster 3"),
    labels_gp = txt_gp,
    width = unit(3, "mm"),
    gp = gpar(fill = c(rep("black", 3)), col = "transparent")),
  which = "row")
lfc_df <- lfc_df[,order(channel_conditions)]
colnames(lfc_df) -> l
lfc_df[protein_list[protein_list %in% rownames(lfc_df)],] -> ht_data

# label the most up- and down-regulated proteins, averaged across all conditions
if (!is.null(mark_top_n)) {
  protein_means <- sort(rowMeans(ht_data, na.rm = TRUE), decreasing = TRUE)
  proteins_to_mark <- c(head(names(protein_means), mark_top_n),
                        tail(names(protein_means), mark_top_n))
}

ha <- rowAnnotation(proteins = anno_mark(
  at = match(proteins_to_mark,rownames(ht_data)),
  labels = proteins_to_mark,
  which = "row",side = "right",
  labels_gp =  gpar(col = "black", fontsize = 5, font="Arial"),
  link_gp = gpar(size = LINE_WIDTH, width = 1,col = "black", lwd = 1.5),
  
  ))


ht_data %>%
  Heatmap(
    name = "log2(fold-change)",
    show_row_names = FALSE,
    top_annotation = ba,
    right_annotation = ha,
    row_title = NULL,
    column_split = sort(channel_conditions),
    column_title = NULL,
    cluster_columns = FALSE,
    show_row_dend = FALSE,
    cluster_column_slices = TRUE,
    show_column_names = FALSE,
    show_column_dend = FALSE,
    heatmap_legend_param = list(
            title = expression("log"[2]*"(FC from M0)"),
            legend_height = unit(0,"cm"),
            legend_width = unit(1.75,"cm"),
            direction = "horizontal",
            title_gp = gpar(col = "black", fontsize = 6, font="Arial"),
            labels_gp = gpar(col = "black", fontsize = 6, font="Arial"),
            grid_width = unit(1, "mm")
          ),
    col = colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))
    ) -> ht


pdf(
  file = paste0(results_name, "lfc_heatmap.pdf"),
  height = 2.1,
  width = 2.7
)

draw(ht, heatmap_legend_side = "bottom")

dev.off()

draw(ht, heatmap_legend_side = "bottom")

ht_data <- ht_data[row_order(ht),] %>% rownames_to_column("protein") 
ht_data |>  write_csv(paste0(results_name, "lfc_data.csv"))

return(ht_data)
}

ht_data <- wp_with_heatmap_labels(
  protein_list,
  c("CD274","IDO1","ISG15","IFIT1","NAMPT","PTGS2","STAT4","TRAFD1","SLAMF7","OAS3","DDX60","MX2","PTPN1","TXNIP","JAK2","CD74", "PTPN2", "STAT1","STAT2", "OAS2", "IFIH1"),
  '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 1_RNAseq, TMT-exp/Panels/heatmap/ifng_heatmap_'
)
ht_data <- wp_with_heatmap_labels(
  # Glycolysis panel = HALLMARK_GLYCOLYSIS (MSigDB 2026.1.Hs), matching the volcano panels.
  msigdb_genes("HALLMARK_GLYCOLYSIS", "H"),
  # mark the 5 most up- and 5 most down-regulated proteins (mean log2FC across all conditions)
  results_name = "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 1_RNAseq, TMT-exp/Panels/heatmap/glycolysis_",
  mark_top_n = 5
)

ht_data
```



## GO enrichment

Fisher-exact GO enrichment for the WP figures (was `run_wp_go_enrichment.R`). Runs the shared
`go_enrich_multi()` engine (`macrophage_figure_setting.R`: clusterProfiler::enricher on MSigDB
2026 GO + simplify) and writes the standard-schema CSVs that the GO-term and correlation-GO
chunks below read. Gene-list selection stays in `wp_downstream_analysis.ipynb`, which exports the
inputs. Each comparison keeps its own experiment-specific background.

Inputs (written by `wp_downstream_analysis.ipynb`):
  <ref_dir>/wp_changing_genes.csv      condition, changing_genes_tested, gene  (targets)
  <ref_dir>/wp_background.csv          gene                                    (WP universe)
  <wp_dir>/rna_vs_wp_go_targets.csv    condition, quadrant, direction, gene    (targets)
  <wp_dir>/rna_vs_wp_go_background.csv condition, gene                         (per-comparison universe)

```{r}
ONTS <- c("BP", "MF", "CC")

ref_dir <- "/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/04_results/goterm_analysis/experiment_specific_background"
wp_dir  <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/01_data_analysis_folders/06_macrophage/macrophage/whole_proteome"



# ---- 2) RNA-vs-WP correlation quadrants (background = per-comparison detected proteome) ----
targets <- readr::read_csv(file.path(wp_dir, "rna_vs_wp_go_targets.csv"), show_col_types = FALSE)
bg      <- readr::read_csv(file.path(wp_dir, "rna_vs_wp_go_background.csv"), show_col_types = FALSE)

groups <- dplyr::distinct(targets, condition, quadrant, direction)
for (i in seq_len(nrow(groups))) {
  g <- groups[i, ]
  tg <- targets |>
    dplyr::filter(condition == g$condition, quadrant == g$quadrant, direction == g$direction) |>
    dplyr::pull(gene) |> unique()
  universe <- bg |> dplyr::filter(condition == g$condition) |> dplyr::pull(gene) |> unique()
  # simplify_cutoff = 0.4 (more aggressive than the 0.5 default): this panel spans 7
  # comparisons, so stronger GO redundancy reduction is warranted to keep terms legible.
  res <- go_enrich_multi(tg, onts = ONTS, background = universe, simplify_cutoff = 0.4)
  out <- file.path(wp_dir, paste0(g$condition, "_", g$quadrant, "_", g$direction, ".csv"))
  readr::write_csv(res, out)
  message("wrote ", basename(out), " (", nrow(res), " rows)")
}
```

## GO term lists
```{r, fig.width=3.5, fig.height = 2}

for (direction in c("up", "down", "up_and_down")) {
  paste0(
    "/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/04_results/goterm_analysis/experiment_specific_background/aggregated_long_goterm_ref_list_",
    direction,
    "_go_term_results.csv"
  ) %>% read_csv(show_col_types = FALSE) %>%
    # standard schema from go_enrich() (GO enrichment chunk above) -> column names used below
    mutate(term = Term, n_genes = n_proteins, p_fdr_bh = adjusted_p_value) -> df
  
  terms_to_plot <- c(
    "mitochondrial translation",
    "ribosomal small subunit biogenesis",
    "ribosomal large subunit biogenesis",
    "rRNA processing",
    "glycolytic process",
    "Cell adhesion",
    "cellular response to lipopolysaccharide",
    "lipid metabolic process",
    "vesicle-mediated transport"
  )
  
  #df %>% filter(class == "biological_process", significant == TRUE) %>% pull(term) %>% unique() %>% head(8)-> terms_to_plot
  df %>%
    # select terms to plot
    filter(toupper(term) %in% toupper(terms_to_plot)) -> plt_df
  #plt_df$term <- factor(plt_df$term, terms_to_plot)
  plt_df$condition <- factor(plt_df$condition, conditions)
  
  plt_df %>%
    filter(n_genes != 0) %>%
    filter(n_genes > 5) %>%
    arrange(term) %>%
    filter(p_fdr_bh < 0.05) %>%
    ggplot(aes(
      x = condition,
      y = str_wrap(term, 20),
      size = n_genes,
      fill = -log10(p_fdr_bh)
    )) +
    geom_point(shape = 21, stroke = POINT_STROKE) +
    #guides(fill = guide_legend(title.position = "right", title.hjust=0.5)) +
    scale_fill_gradientn(
      colors = c("white", "#5CBED7"),
      #rescaler = ~ scales::rescale_mid(.x, mid = -log10(0.05)),
      breaks = c(round(-log10(0.05), 2), 5),
      limits = c(round(-log10(0.05), 2), 5),
      na.value = "#5CBED7"
    ) +
    my_theme() +
    labs(
      x = NULL,
      y = NULL,
      fill = expression("-log"[10] * "(P"["adj"] * ")"),
      size = "Gene count"
    ) +
    #guides(size = guide_legend(title.position = "right", title.hjust=0.5)) +
    theme(
      panel.grid.major = element_line(colour = "grey90", linewidth = LINE_WIDTH),
      legend.location = "bottom",
      legend.position = "bottom",
      legend.text = element_text(
        family = "Arial",
        size = 6,
        margin = margin(0, 0, 0, 0)
      ),
      legend.key.size = unit(4, "mm"),
      legend.spacing = unit(0, "mm"),
      legend.margin = margin(l = 0, b = 0),
      legend.title = element_text(family = "Arial", size = 6)
    ) -> plt
  print(plt)
  
  save_plot(
    paste0(
      "/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/04_results/goterm_analysis/",
      direction,
      "_0.05_fdr_min_5_genes"
    ),
    width = 3.8,
    height = 2
  )
}
```


```{r, fig.width=3}

library(gridExtra)
library(gtable)
library(grid)

# Fig 1H layout: rows = conditions (coloured like the volcano panels), columns = GO category,
# split into up- and down-regulated protein counts under two spanning header bands.
cond_levels <- c("TLR1-2", "TLR3", "TLR4", "TLR7", "TLR8", "STING")
cat_levels  <- c("Electron transport chain", "Glycolysis",
                 "Mitochondrial", "Mitochondrial translation")
cat_labels  <- c("ETC", "glycolysis", "mt", "mt-\ntranslation")


go_terms_fn <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 1_RNAseq, TMT-exp/Panels/volcano_plot/2026_ontology/volcano_data_long_format.csv"
volcano_data <- read_csv(go_terms_fn, show_col_types = FALSE)

counts <- volcano_data %>%
  mutate(Mitochondrial = Mitochondrial & !`Mitochondrial translation`,
         Glycolysis     = Glycolysis    & !`Electron transport chain`) %>%
  select(Regulation, all_of(cat_levels), condition) %>%
  pivot_longer(all_of(cat_levels), names_to = "name") %>%
  filter(value, Regulation %in% c("Significant Up", "Significant Down"),
         condition %in% cond_levels) %>%
  count(Regulation, name, condition)
counts %>% arrange(Regulation, name, condition) %>% write_csv("categories_to_plot.csv")

# conditions x categories count matrix for one direction (0 where a category had no hits)
fill_block <- function(reg) {
  m <- matrix(0L, length(cond_levels), length(cat_levels),
              dimnames = list(cond_levels, cat_levels))
  d <- dplyr::filter(counts, Regulation == reg)
  if (nrow(d)) m[cbind(match(d$condition, cond_levels), match(d$name, cat_levels))] <- d$n
  m
}
disp <- as.data.frame(cbind(fill_block("Significant Up"), fill_block("Significant Down")))
colnames(disp) <- rep(cat_labels, 2)

# white bordered cells, Arial, multi-line sub-headers
th <- ttheme_default(
  base_size = 6,
  padding = unit(c(3, 2.4), "mm"),
  core    = list(fg_params = list(fontfamily = "Arial"),
                 bg_params = list(fill = "white", col = "grey30", lwd = 0.5)),
  colhead = list(fg_params = list(fontfamily = "Arial", fontface = "plain"),
                 bg_params = list(fill = "white", col = "grey30", lwd = 0.5)),
  rowhead = list(fg_params = list(fontfamily = "Arial", fontface = "plain")))
g <- tableGrob(disp, theme = th)

# colour each condition's row label like the volcano panels (unique names so the added
# grobs aren't dropped as duplicates of tableGrob's own "rowhead"/"core" grobs)
for (i in seq_along(cond_levels)) {
  g <- gtable_add_grob(g, rectGrob(gp = gpar(fill = cols[[cond_levels[i]]], col = "grey30", lwd = 0.5)),
                       t = 1 + i, l = 1, z = 0, name = paste0("rowlbl-", i))
}

# spanning header band over each category group
ncat <- length(cat_levels)
g <- gtable_add_rows(g, heights = unit(5, "mm"), pos = 0)
add_group <- function(g, label, l, r, fill, tag) {
  g <- gtable_add_grob(g, rectGrob(gp = gpar(fill = fill, col = "grey30", lwd = 0.5)),
                       t = 1, b = 1, l = l, r = r, z = 0, name = paste0("band-bg-", tag))
  gtable_add_grob(g, textGrob(label, gp = gpar(fontsize = 6, fontfamily = "Arial")),
                  t = 1, b = 1, l = l, r = r, name = paste0("band-fg-", tag))
}
g <- add_group(g, "# of upregulated proteins",   2,        1 + ncat,     "#EAA9A9", "up")
g <- add_group(g, "# of downregulated proteins", 2 + ncat, 1 + 2 * ncat, "#A9C7E8", "down")

pdf("/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 1_RNAseq, TMT-exp/Panels/volcano_plot/2026_ontology/go_term_list_table.pdf",
    height = 1.6, width = 4.2)
grid.arrange(g)
dev.off()

grid.arrange(g)
```




```{r, fig.height=2, fig.width=3}
library(patchwork)
library(ComplexUpset)
source('/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/01_data_analysis_folders/06_macrophage/macrophage/macrophage_figure_setting.R')
# standard schema from go_enrich() (GO enrichment chunk above)
'/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/01_data_analysis_folders/06_macrophage/macrophage/whole_proteome/STING_teal-yellow_Up.csv' %>%
  read_csv(show_col_types = FALSE) %>%
  filter(adjusted_p_value < 0.05, n_proteins > 5, ontology != "CC") %>%
  arrange(adjusted_p_value) %>%
  go_dotplot(facet = "ontology", size_range = c(1, 4))

save_plot(
  paste0("/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/01_data_analysis_folders/06_macrophage/macrophage/whole_proteome/correlation_go_term_TLR4_up.svg"), 3, 2
)
```

```{r}
dir <- '/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/04_results/wp_vs_rnaseq/'

conditions <- c("TLR1-2", "TLR3", "TLR4", "TLR7", "TLR8", "TLR9", "STING")
# build sets iteratively
sets <- list()
for(cond in conditions){
  if (cond == "TLR9"){
    next
  }
  sets[[cond]] <- read_csv(paste0(dir, cond, " vs. M0.csv")) %>%
    filter(expression %in% c("yellow", "teal"), logFC_wp > 0) %>% pull(uniprot)
}

# transform to binary data frame
all_elements <- unique(unlist(sets))
df <- as.data.frame(sapply(sets, function(set) {
  all_elements %in% set
}))
```


```{r, fig.height=2.2, fig.width=2.0}

# define plot aesthetics
upset_plot_theme <-
  theme_classic() +
  theme(
    axis.text.y = TEXT_ELEMENT,
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_blank(),
    axis.title.y = TEXT_ELEMENT,
    panel.grid = element_blank(),
    axis.line.x = AXIS_LINE,
    axis.line.y = AXIS_LINE
  )

upset(
  df,
  sort_sets = FALSE,
  intersect = rev(names(sets)),
  set_sizes = FALSE,
  n_intersections = 8,
  matrix = (intersection_matrix(
    geom = geom_point(
      size = 1
    ),
    segment = geom_segment(
      linewidth = LINE_WIDTH
    )
  )),
  annotations = list(
    "intersection size" = intersection_size(
      counts = TRUE,
      color = "black",
      size = LINE_WIDTH,
      show.legend = FALSE,
      width = 0.65,
      fill = cols["TLR8"],
      text = list(size = FONT_SIZE_MM, color = "black", family = FONT_FAMILY, vjust = -0.1, check_overlap = TRUE)
    ) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.2)))
  ),
  base_annotations = list(),
  stripes = "white",
  min_size = 15,
  themes = list(
    `intersection size` = upset_plot_theme + theme(axis.ticks.y = AXIS_LINE),
    intersections_matrix = upset_plot_theme + theme(axis.line.x = element_blank(), axis.line.y = element_blank(), axis.title.y = element_blank())
  )
) + patchwork::plot_layout(heights = c(1, 0.7))
ggsave(paste0(dir, "upset_plot.svg"), height = 2.0, width = 1.9)
```


## Correlation GO dotplot
```{r, fig.height=5, fig.width=4.2}
dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/Figure 2_Unenriched proteomics analysis/Panels/wp_vs_rnaseq_go_term/'
wp_dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/01_data_analysis_folders/06_macrophage/macrophage/whole_proteome'

# Aggregate the per-comparison GO results (written by the GO enrichment chunk above) into the BP
# correlation frame, then write it for downstream/manuscript use. Was cell 30 of
# wp_downstream_analysis.ipynb; moved here so it tracks the current go_enrich() schema
# (adjusted_p_value / n_proteins / ontology) instead of a stale notebook artifact.
go_bp <- list()
for (cond in conditions) {
  for (dir_lvl in c("Up", "Down", "Both")) {
    for (clr in c("violet", "teal-yellow")) {
      fn <- file.path(wp_dir, paste0(cond, "_", clr, "_", dir_lvl, ".csv"))
      if (!file.exists(fn)) next
      df <- read_csv(fn, show_col_types = FALSE) %>%
        mutate(condition = cond, direction = dir_lvl, color = clr) %>%
        filter(adjusted_p_value < 0.05, n_proteins > 5, ontology == "BP")
      go_bp[[length(go_bp) + 1]] <- df
    }
    print(cond)
    print(length(df))
  }
}
go_bp <- bind_rows(go_bp)
write_csv(go_bp, paste0(dir, "rna_vs_wp_correlation_go_term_analysis_BP.csv"))

plotted <- go_bp %>%
  filter(adjusted_p_value < 0.01, n_proteins > 5, direction != "Both", color == "teal-yellow", condition != "TLR9") %>%
  mutate(condition = factor(condition, levels = conditions), term = str_wrap(Term, 50))

# Diagonal "staircase" ordering (classic heatmap look): assign each term to the
# condition/direction where it is most enriched (its peak -log10 FDR), then order terms by
# that condition so TLR1-2-specific terms sit at the top and STING-specific ones at the
# bottom, with within-condition ties broken by significance.
term_levels <- plotted %>%
  group_by(term) %>%
  slice_max(neg_log10_fdr, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(desc(condition), desc(direction), desc(neg_log10_fdr)) %>%
  pull(term)

plotted %>%
  mutate(term = factor(term, levels = term_levels)) %>%
  ggplot(aes(x = direction, y = term, size = n_proteins, fill = neg_log10_fdr)) +
  geom_point(shape = 21, color = "black", stroke = LINE_WIDTH) + 
  facet_grid(rows = vars(color), cols = vars(condition), scales = "free") + 
  my_theme() +
  theme(
    panel.grid.major = element_line(color = "grey90", size = LINE_WIDTH),
    panel.spacing = unit(0, "mm"),
    legend.text = element_text(family = FONT_FAMILY, size = FONT_SIZE, color = "black", margin = margin()),
    strip.background = element_blank(),
    strip.clip = "off",
    axis.text.y = element_text(size = 5, family = FONT_FAMILY, color = "black"),
    axis.text.x = element_text(size = 5, family = FONT_FAMILY, color = "black", angle = 45, vjust =1, hjust = 1),
    strip.text = element_text(size = 5.5, color = "black", family= FONT_FAMILY)
    ) + 
  labs(size = "proteins", y = NULL) + 
  scale_fill_gradient2(low = "#edf8fb", mid = "#8c96c6", high = "#810f7c") +
  scale_size_continuous(range = c(1, 4))
save_plot(paste0(dir, "go_term_correlation_BP"), width = 4.2, height = 5)

go_bp$condition
dir
```








# Binned half life data



```{r, fig.width=6, fig.height=2}
dir <-  '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/Figure 2_Unenriched proteomics analysis/Panels/wp_vs_rnaseq'
palette <-   c(
  "96+ h" = "#721f81",
  "48-96 h" = "#b73779",
  "24-48 h" = "#f1605d",
  "0-24 h" = "#feb078"
)

immune_relevance<- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/papers/chemical proteomics/Cell_2020/immune_relevant_all.csv' |> read_csv(show_col_types = FALSE) |> select(z_sum, symbol) |>mutate( binned_immune_relevance = case_when(
        z_sum < 12 ~ "not immune relevant (0-12)",
        z_sum < 24 ~ "immune relevant (12-24)",
        TRUE ~ "highly immune relevant (24+)"
      ))
condition <- "TLR3"
df <- NULL
for (cond in names(cols)) {
  if (cond %in% c("M0", "TLR9")) {
    next
  }
  print(cond)
  df <- paste0(dir, "/", cond, " vs. M0_annotated.csv") %>%
    read_csv(show_col_types = FALSE) %>%
    mutate(
      direction = case_when(logFC_wp > 0 ~ "upregulated", TRUE ~ "downregulated"),
      binned_half_life = case_when(
        mean_monocyte_half_life < 24 ~ "0-24 h",
        mean_monocyte_half_life < 48 ~ "24-48 h",
        mean_monocyte_half_life < 96 ~ "48-96 h",
        TRUE  ~ "96+ h"
      ),
      expression = factor(expression, levels = c("teal", "yellow", "dark_grey", "violet")),
      condition = cond
    ) %>%
    select(expression,
           direction,
           mean_monocyte_half_life,
           binned_half_life,
           condition) %>%
    filter(expression != "light_grey") %>%
    drop_na() %>%
    rbind(df)
}

df %>%
  ggplot(aes(y = expression, fill = binned_half_life)) + geom_bar(position = "fill",
                                                                  color = "black",
                                                                  size = LINE_WIDTH) +
  my_theme() +
  scale_fill_manual(values = palette) +
  facet_grid(cols = vars(condition), rows = vars(direction)) +
  scale_x_continuous(expand = expansion(0, 0), labels = scales::percent_format(suffix = "")) +
  scale_y_discrete(expand = expansion(0, 0)) +
  labs(y = "quadrant", x = "% of proteins", fill = "") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE
    ),
    strip.clip = "off",
    legend.position = "bottom",
    legend.key.size = unit(3, "mm"),
    legend.text = element_text(size = FONT_SIZE, color = "black", family = FONT_FAMILY, margin = margin()),
    legend.margin = margin(),
    legend.box.margin = margin(),
    panel.spacing = unit(4, "mm")
  ) + labs(y = NULL)

ggsave(paste0(dir, "half_life_by_quadrant.svg"),
       width = 6,
       height = 2)

```

```{r, fig.width=4, fig.height=3}
df %>%
  ggplot(aes(x = expression, fill = binned_half_life)) + geom_bar(
                                                                  color = "black",
                                                                  size = LINE_WIDTH) +
  my_theme() +
  scale_fill_manual(values = palette) +
  facet_grid(cols = vars(condition), rows = vars(direction)) +
  #scale_x_continuous(expand = expansion(0, 0), labels = scales::percent_format(suffix = "")) +
  #scale_y_discrete(expand = expansion(0, 0)) +
  labs(y = "quadrant", x = "% of proteins", fill = "") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE
    ),
    strip.clip = "off",
    legend.position = "bottom",
    legend.key.size = unit(3, "mm"),
    legend.text = element_text(size = FONT_SIZE, color = "black", family = FONT_FAMILY, margin = margin()),
    legend.margin = margin(),
    legend.box.margin = margin(),
    panel.spacing = unit(0, "mm"),
    axis.text.x = element_text(
        color="black",angle=45, size=FONT_SIZE, family="Arial", vjust = 1, hjust=1)
  ) + labs(x = NULL, y = "number of proteins")
```


```{r, fig.width = 2, fig.height=2}
immune_relevance<- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/papers/chemical proteomics/Cell_2020/immune_relevant_all.csv' |> read_csv(show_col_types = FALSE) |> select(z_sum, symbol) |>mutate( binned_immune_relevance = case_when(
        z_sum < 12 ~ "not immune relevant (0-12)",
        z_sum < 24 ~ "immune relevant (12-24)",
        TRUE ~ "highly immune relevant (24+)"
      ))
```




## Binned immune relevance

```{r}
immune_relevance<- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/papers/chemical proteomics/Cell_2020/immune_relevant_all.csv' |>
  read_csv()

immune_relevance |> 
  select(symbol, z_sum) |> 
  ggplot(aes(x = z_sum)) + 
  geom_histogram()
```

```{r, fig.width=4, fig.height=2}
library(dplyr)
library(ggplot2)
immune_relevance<- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/papers/chemical proteomics/Cell_2020/immune_relevant_all.csv' |>
  read_csv()
# Calculate percentiles
percentiles <- immune_relevance |>
  select(symbol, z_sum) |>
  mutate(z_sum = replace_na(z_sum, 0)) |>
  summarise(
    p50 = quantile(z_sum, 0.50),
    p75 = quantile(z_sum, 0.75),
    p90 = quantile(z_sum, 0.90),
    p99 = quantile(z_sum, 0.95)
  )

# Print the table
percentiles |>
  tidyr::pivot_longer(everything(), names_to = "percentile", values_to = "value") |>
  mutate(percentile = recode(percentile,
    p50 = "50th", p75 = "75th", p90 = "90th", p99 = "95th"
  )) |>
  print()

# Plot histogram with percentile lines
immune_relevance |>
  select(symbol, z_sum) |>
  mutate(z_sum = replace_na(z_sum, 0)) |>
  ggplot(aes(x = z_sum)) +
  geom_histogram() +
  geom_vline(
    data = tidyr::pivot_longer(percentiles, everything(),
                               names_to = "percentile", values_to = "value"),
    aes(xintercept = value, color = percentile, linetype = percentile),
    linewidth = 0.8
  ) +
  scale_color_manual(
    values = c(p50 = "#E41A1C", p75 = "#FF7F00", p90 = "#984EA3", p99 = "#377EB8"),
    labels = c(p50 = "50th", p75 = "75th", p90 = "90th", p99 = "95th")
  ) +
  scale_linetype_manual(
    values = c(p50 = "solid", p75 = "dashed", p90 = "dotdash", p99 = "dotted"),
    labels = c(p50 = "50th", p75 = "75th", p90 = "90th", p99 = "95th")
  ) +
  labs(color = "Percentile", linetype = "Percentile") +
  theme_minimal() + 
  labs(y = "genes", x= "summed Z-score")
ggsave(paste0(dir, "/immune relevance_by_quadrant.pdf"),
       width = 4,
       height = 2)
```



```{r, fig.width=6, fig.height=2}
dir <-  '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/Figure 2_Unenriched proteomics analysis/Panels/wp_vs_rnaseq'
palette <-   c(
  "not immune enriched\n(0-75th percentile)" = "#5ec962",
  "moderately immune enriched\n(75th-90th percentile)" = "#21918c",
  "immune enriched\n(90th-95th percentile)" = "#3b528b",
  "highly immune enriched\n(95th+ percentile)" = "#440154"
)

immune_relevance<- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/papers/chemical proteomics/Cell_2020/immune_relevant_all.csv' |> read_csv(show_col_types = FALSE) |> select(z_sum, symbol) |>mutate( binned_immune_relevance = case_when(
  is.na(z_sum) ~ "not immune enriched\n(0-75th percentile)",
        z_sum < 4.2 ~ "not immune enriched\n(0-75th percentile)",
        z_sum < 11.5 ~ "moderately immune enriched\n(75th-90th percentile)",
        z_sum < 33.3 ~ "immune enriched\n(90th-95th percentile)",
        TRUE ~ "highly immune enriched\n(95th+ percentile)"
      )) |> 
  rename(protein = symbol)
condition <- "TLR3"
df <- NULL
for (cond in names(cols)) {
  if (cond %in% c("M0", "TLR9")) {
    next
  }
  print(cond)
  df <- paste0(dir, "/", cond, " vs. M0_annotated.csv") %>%
    read_csv(show_col_types = FALSE) %>% 
    mutate(
      direction = case_when(logFC_wp > 0 ~ "upregulated", TRUE ~ "downregulated"),
      expression = factor(expression, levels = c("teal", "yellow", "dark_grey", "violet")),
      condition = cond
    ) %>%
    left_join(y = immune_relevance, by = "protein") |> 
    select(expression,
           direction,
           binned_immune_relevance,
           condition) %>%
    filter(expression != "light_grey") %>%
    drop_na() %>%
    rbind(df)
}

df %>%
  ggplot(aes(y = expression, fill = binned_immune_relevance)) + geom_bar(position = "fill",
                                                                  color = "black",
                                                                  size = LINE_WIDTH) +
  my_theme() +
  scale_fill_manual(values = palette) +
  facet_grid(cols = vars(condition), rows =vars(direction)) +
  scale_x_continuous(expand = expansion(0, 0), labels = scales::percent_format(suffix = "")) +
  scale_y_discrete(expand = expansion(0, 0)) +
  labs(y = "quadrant", x = "% of proteins", fill = "") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE
    ),
    strip.clip = "off",
    legend.position = "bottom",
    legend.key.size = unit(3, "mm"),
    legend.text = element_text(size = FONT_SIZE, color = "black", family = FONT_FAMILY, margin = margin()),
    legend.margin = margin(),
    legend.box.margin = margin(),
    panel.spacing = unit(4, "mm")
  ) + labs(y = NULL)

ggsave(paste0(dir, "/immune relevance_by_quadrant.pdf"),
       width = 6,
       height = 1.75)

df

```

```{r}
immune_relevance<- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/papers/chemical proteomics/Cell_2020/immune_relevant_all.csv' |> read_csv(show_col_types = FALSE)

immune_relevance$`immune phenotype? (OMIM)`
```


```{r, fig.width=6, fig.height=2}
dir <-  '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/Figure 2_Unenriched proteomics analysis/Panels/wp_vs_rnaseq'
palette <-   c(
  "not immune enriched\n(0-75th percentile)" = "#5ec962",
  "moderately immune enriched\n(75th-90th percentile)" = "#21918c",
  "immune enriched\n(90th-95th percentile)" = "#3b528b",
  "highly immune enriched\n(95th+ percentile)" = "#440154"
)

immune_relevance<- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/papers/chemical proteomics/Cell_2020/immune_relevant_all.csv' |> read_csv(show_col_types = FALSE) |> 
  select(`immune phenotype? (OMIM)`, symbol) |> 
  rename(protein = symbol)
condition <- "TLR3"
df <- NULL
for (cond in names(cols)) {
  if (cond %in% c("M0", "TLR9")) {
    next
  }
  print(cond)
  df <- paste0(dir, "/", cond, " vs. M0_annotated.csv") %>%
    read_csv(show_col_types = FALSE) %>% 
    mutate(
      direction = case_when(logFC_wp > 0 ~ "upregulated", TRUE ~ "downregulated"),
      expression = factor(expression, levels = c("teal", "yellow", "dark_grey", "violet")),
      condition = cond
    ) %>%
    left_join(y = immune_relevance, by = "protein") |> 
    select(expression,
           direction,
           `immune phenotype? (OMIM)`,
           condition) %>%
    filter(expression != "light_grey") %>%
    drop_na() %>%
    rbind(df)
}

df %>%
  ggplot(aes(y = expression, fill = `immune phenotype? (OMIM)`)) + geom_bar(position = "fill",
                                                                  color = "black",
                                                                  size = LINE_WIDTH) +
  my_theme() +
  facet_grid(cols = vars(condition), rows =vars(direction)) +
  scale_x_continuous(expand = expansion(0, 0), labels = scales::percent_format(suffix = "")) +
  scale_y_discrete(expand = expansion(0, 0)) +
  labs(y = "quadrant", x = "% of proteins", fill = "") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE
    ),
    strip.clip = "off",
    legend.position = "bottom",
    legend.key.size = unit(3, "mm"),
    legend.text = element_text(size = FONT_SIZE, color = "black", family = FONT_FAMILY, margin = margin()),
    legend.margin = margin(),
    legend.box.margin = margin(),
    panel.spacing = unit(4, "mm")
  ) + labs(y = NULL, fill = "OMIM phenotype")

ggsave(paste0(dir, "/immune relevance_by_quadrant.pdf"),
       width = 6,
       height = 1.75)

df

```

```{r, fig.width=4, fig.height=3}
df %>%
  ggplot(aes(x = expression, fill = binned_half_life)) + geom_bar(
                                                                  color = "black",
                                                                  size = LINE_WIDTH) +
  my_theme() +
  scale_fill_manual(values = palette) +
  facet_grid(cols = vars(condition), rows = vars(direction)) +
  #scale_x_continuous(expand = expansion(0, 0), labels = scales::percent_format(suffix = "")) +
  #scale_y_discrete(expand = expansion(0, 0)) +
  labs(y = "quadrant", x = "% of proteins", fill = "") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE
    ),
    strip.clip = "off",
    legend.position = "bottom",
    legend.key.size = unit(3, "mm"),
    legend.text = element_text(size = FONT_SIZE, color = "black", family = FONT_FAMILY, margin = margin()),
    legend.margin = margin(),
    legend.box.margin = margin(),
    panel.spacing = unit(0, "mm"),
    axis.text.x = element_text(
        color="black",angle=45, size=FONT_SIZE, family="Arial", vjust = 1, hjust=1)
  ) + labs(x = NULL, y = "number of proteins")
```

# Ferry data

```{r, fig.width=2, fig.height=2}
volcano_data <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/census-out-processing/long_volcano_data.csv' |> 
  read_csv(show_col_types = FALSE)
```

## Volcano plots

```{r, fig.width = 2, fig.height=3}
volcano_data <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/census-out-processing/long_volcano_data.csv' |> 
  read_csv(show_col_types = FALSE)

options(ggrepel.max.overlaps = Inf) # for crowded volcano plots

endogenous_color_code <- c(
  "FERRY" = "#5ebed7",
  "RBP" = "#496ab2",
  "Ribosomal" = "#267e41",
  "Other" = "#ababab"
)
endogenous_volcano_dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/endogenous_pulldown_analysis/2_plus_reps_2_plus_peps/volcano_plots_updated_cutoff/'

for (exp_comparison in unique(volcano_data$comparison)){
  y_max <- NULL
  if (exp_comparison %in% c("M0-Fy2 vs. M0-IgG", "LPS-Fy2 vs. LPS-IgG")){
    y_max <- 20.6501079122941
  }
  # LPS-vs-M0 cross-condition comparisons use a 2x FC cutoff; the vs-IgG
  # enrichments stay at 1.5x. This drives both the recoded Regulation column
  # (point coloring / count boxes) and the dashed guide lines below.
  fc_cutoff_val <- if (exp_comparison %in% c("LPS-Fy2 vs. M0-Fy2", "LPS-Tbck vs. M0-Tbck")) {
    log2(2)
  } else {
    log2(1.5)
  }
  print(exp_comparison)
  condition = str_split(exp_comparison, " vs. ")[[1]][1]
  control = str_split(exp_comparison, " vs. ")[[1]][2]
  plt_data <- volcano_data |>
    filter(comparison == exp_comparison) |>
    mutate(
      group = category,
      log2_FC = ifelse(is.infinite(log2_FC), 7, log2_FC),
      # recode Regulation in R at fc_cutoff_val (overrides the CSV's 1.5x-based
      # values); strings must match what make_condition_plot_multi checks for
      Regulation = case_when(
        log2_FC >  fc_cutoff_val & `-log10_pval` > -log10(0.05) ~ "Significant Up",
        log2_FC < -fc_cutoff_val & `-log10_pval` > -log10(0.05) ~ "Significant Down",
        TRUE ~ "Stable"
      )
    )

  make_condition_plot_multi(
    condition = condition,
    volcano_data = plt_data,
    control_condition = control,
    color_code = endogenous_color_code,
    plot_directory = endogenous_volcano_dir,
    x_axis_word = "pulldown",
    highlight_proteins = c("TBCK", "CRYZL1", "C12orf4", "PPP1R21", "GATD1", "RAB5C", "DISC1"),
    y_max = y_max,
    fc_cutoff = fc_cutoff_val
  )
}


```

```{r, fig.width=2, fig.height=4}
correlation_plot <-
  function(combined_df,
    timepoints,
           min,
           max,
           correlation_palette,
           num_top = 6,
           x_lab,
           y_lab,
           drop = NULL,
           top = c()) {
    # constants for defining position of margins
    margin_factor <- 0.85
    margin_position <- min * margin_factor
    missing_point_value <- (min + margin_position) / 2

    log_threshold <- -log10(0.05)


    wide_df <- combined_df %>%
      filter(comparison %in% unname(timepoints) )|> 
      mutate(time_point = recode(comparison, !!!setNames(names(timepoints), timepoints)), log10_pval = `-log10_pval`, alias = protein) |> 
      select(alias, log10_pval, log2_FC, time_point) |> 
      pivot_wider(
        names_from = time_point,
        values_from = c(log2_FC, log10_pval)
      )
    
    print(colnames(wide_df))

    
    combined_df <- wide_df %>%
      mutate(
        significance = case_when(
          (log10_pval_x > log_threshold) &
            (log10_pval_y > log_threshold) ~ names(correlation_palette[4])[1],
          (log10_pval_y > log_threshold) ~ names(correlation_palette[2])[1],
          (log10_pval_x > log_threshold) ~ names(correlation_palette[1])[1],
          TRUE ~ names(correlation_palette[3])[1]
        )
      ) %>%
      mutate(
        #significance = factor(significance, levels = names(correlation_palette)),
        dist_from_orig = sqrt(log2_FC_x^2 + log2_FC_y^2),
        distance_from_identity = abs(log2_FC_x - log2_FC_y)
      ) 

    # calculate pearson's correlation from complete cases
    pearson_r <- cor(combined_df$log2_FC_x, combined_df$log2_FC_y, use = "complete.obs") %>%
      round(digits = 2)


    print(max(combined_df$log2_FC_x, na.rm = TRUE))
    print(max(combined_df$log2_FC_y, na.rm = TRUE))
    print(min(combined_df$log2_FC_x, na.rm = TRUE))
    print(min(combined_df$log2_FC_y, na.rm = TRUE))

    combined_df <- combined_df %>%
      # replace missing values with values on the margins
      replace_na(list(log2_FC_x = missing_point_value, log2_FC_y = missing_point_value))

    # pull list of proteins to give text labels
    significant_no_missing <- combined_df %>%
      filter(
        significance != "Not significant",
        log2_FC_x != missing_point_value,
        log2_FC_y != missing_point_value
      )
    label_data <- rbind(
      # top by distance from identity
      significant_no_missing %>%
        arrange(distance_from_identity) %>%
        tail(num_top),
      # top by distance from origin
      significant_no_missing %>%
        arrange(dist_from_orig) %>%
        tail(num_top),
      # top only quantified at x
      combined_df %>%
        arrange(abs(log2_FC_x)) %>%
        filter(
          significance != "Not significant",
          (log2_FC_y == missing_point_value)
        ) %>%
        tail(num_top / 2),
      # top only quantified at y
      combined_df %>%
        arrange(abs(log2_FC_y)) %>%
        filter(
          significance != "Not significant",
          (log2_FC_x == missing_point_value)
        ) %>%
        tail(num_top / 2)
    ) %>% distinct()
    
    if (length(top) > 0){
      label_data <- combined_df %>% filter(significance != "Not significant", alias %in% top)
    }
    
    print(top[!top %in% unique(label_data$alias)])
    print("CAVIN2" %in% combined_df$alias)
    print("SDPR" %in% combined_df$alias)

    plot <- combined_df %>%
      # sort so labeled points are stacked on top
      mutate(z_order = alias %in% unique(label_data$alias)) %>%
      arrange(z_order) %>%
      ggplot(aes(
        x = log2_FC_x,
        y = log2_FC_y,
        fill = significance,
        label = alias
      )) +
      # margin lines
      geom_hline(
        yintercept = margin_position,
        size = LINE_WIDTH,
        color = "black"
      ) +
      geom_vline(
        xintercept = margin_position,
        size = LINE_WIDTH,
        color = "black"
      ) +
      # origin lines
      geom_hline(
        yintercept = 0,
        color = "grey90",
        size = LINE_WIDTH
      ) +
      geom_vline(
        xintercept = 0,
        color = "grey90",
        size = LINE_WIDTH
      ) +
      # diagonal boundary lines
      geom_segment(
        data = data.frame(),
        aes(
          x = c(margin_position, margin_position, margin_position + 1),
          y = c(margin_position + 1, margin_position, margin_position),
          xend = c(max - 1, max, max),
          yend = c(max, max, max - 1),
          linetype = c("dashed", "solid", "dashed")
        ),
        size = LINE_WIDTH,
        inherit.aes = FALSE,
        show.legend = FALSE,
        color = "black"
      ) +
      # scatterplot
      geom_point(
        shape = 21,
        stroke = 0,
        alpha = 0.55,
        size = 1.4,
      ) +
      # protein labels
      geom_text_repel(
        data = label_data,
        size = 1.94028,
        family = FONT_FAMILY,
        min.segment.length = 0,
        #force = 3,
        segment.size = LINE_WIDTH,
        max.overlaps = Inf,
        #nudge_y = -2,
        #nudge_x = 2,
        show.legend = FALSE#,
        #segment.curvature = -0.1,
      #segment.ncp = 3,
      #segment.angle = 45
      ) +
      my_theme() +
      theme(
        aspect.ratio = 1,
        # 20 lines of legend formatting
        legend.position = "top",
        legend.text = element_text(
          size = 6,
          color = "black",
          family = "Arial",
          margin = margin(l = 0)
        ),
        legend.box.background = element_blank(),
        legend.justification.top = "top",
        legend.justification = "center",
        legend.box.just = "center",
        legend.key = element_blank(),
        legend.margin = margin(l = -5, b = -5.5),
        legend.box.margin = margin(l = -5, b = -0.8),
        legend.key.spacing.y = unit(-0.9, "cm"),
        legend.key.spacing.x = unit(-0.1, "cm")
      ) +
      guides(fill = guide_legend(nrow = 2, override.aes = list(alpha = 1))) +
      labs(
        x = x_lab,
        y = y_lab
      ) +
      scale_fill_manual(values = correlation_palette, name = NULL) +
      scale_x_continuous(expand = expansion(mult = 0), limits = c(min, max)) +
      scale_y_continuous(expand = expansion(mult = 0), limits = c(min, max)) +
      scale_linetype_manual(values = c("dashed" = "dashed", "solid" = "solid")) +
      geom_text(
        mapping = aes(
          x = min * 0.6,
          y = max * 0.85,
        ),
        label = paste("R = ", as.character(pearson_r)),
        inherit.aes = FALSE,
        data = data.frame(),
        size = 1.94028,
        family = FONT_FAMILY
      )

    print(plot)
  }



not_significant <- "grey80"
correlation_min <- -3.5
correlation_max <- 8
correlation_plot(
  volcano_data,
  timepoints = c(
    "x" = "LPS-Tbck vs. LPS-IgG", # x is liver
    "y" = "LPS-Fy2 vs. LPS-IgG" # y is fat
  ),
  min = correlation_min, max = correlation_max,
  correlation_palette = c(
    "Tbck p < 0.05" = "#d6e5b3",
    "Fy2 p < 0.05" = "#75c489",
    "Not significant" = not_significant,
    "Both p < 0.05" = "#5d8f70"
  ),
  num_top = 6,
  x_lab = expression("log"[2] * "(LPS-Tbck/LPS-IgG)"),
  y_lab = expression("iWAT, log"[2] * "(LPS-Fy2/LPS-IgG)"),
  drop = "",
  top = c("TBCK", "CRYZL1", "C12orf4", "PPP1R21", "GATD1",  "KPNB1", "FCER1G", "TRIM21")
)

ggsave('/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/endogenous_pulldown_analysis/2_plus_reps_2_plus_peps/correlation_plot/LPS pulldown.png', width= 2, height =4)

correlation_min <- -4.2
correlation_max <- 8.2
correlation_plot(
  volcano_data |> mutate(log2_FC = ifelse(is.infinite(log2_FC), 7.865, log2_FC)),
  timepoints = c(
    "x" = "M0-Tbck vs. M0-IgG", # x is liver
    "y" = "M0-Fy2 vs. M0-IgG" # y is fat
  ),
  min = correlation_min, max = correlation_max,
  correlation_palette = c(
    "Tbck p < 0.05" = "#d6e5b3",
    "Fy2 p < 0.05" = "#75c489",
    "Not significant" = not_significant,
    "Both p < 0.05" = "#5d8f70"
  ),
  num_top = 6,
  x_lab = expression("log"[2] * "(M0-Tbck/M0-IgG)"),
  y_lab = expression("iWAT, log"[2] * "(M0-Fy2/M0-IgG)"),
  drop = "",
  top = c("TBCK", "CRYZL1", "C12orf4", "PPP1R21", "GATD1",  "KPNB1", "FCER1G", "TRIM21", "LRRFIP2", "CRYBG1")
)
volcano_data |> filter(comparison %in% unname(c(
    "x" = "M0-Tbck vs. M0-IgG", 
    "y" = "M0-Fy2 vs. M0-IgG" ))) |> 
    arrange(-log2_FC)
ggsave('/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/endogenous_pulldown_analysis/2_plus_reps_2_plus_peps/correlation_plot/M0 pulldown.png', width= 2, height =4)

correlation_min <- -3.2
correlation_max <- 5.5
correlation_plot(
  volcano_data |> mutate(log2_FC = ifelse(is.infinite(log2_FC), 7.865, log2_FC)),
  timepoints = c(
    "x" = "LPS-Tbck vs. M0-Tbck", # x is liver
    "y" = "LPS-Fy2 vs. M0-Fy2" # y is fat
  ),
  min = correlation_min, max = correlation_max,
  correlation_palette = c(
    "Tbck p < 0.05" = "#d6e5b3",
    "Fy2 p < 0.05" = "#75c489",
    "Not significant" = not_significant,
    "Both p < 0.05" = "#5d8f70"
  ),
  num_top = 6,
  x_lab = expression("log"[2] * "(LPS-Tbck/M0-Tbck)"),
  y_lab = expression("log"[2] * "(LPS-Fy2/M0-Tbck)"),
  drop = "",
  top = c("TBCK", "CRYZL1", "C12orf4", "PPP1R21", "GATD1")
)
volcano_data |> filter(comparison %in% unname(c(
    "x" = "LPS-Tbck vs. M0-Tbck", 
    "y" = "LPS-Fy2 vs. M0-Fy2" ))) |> 
    arrange(log2_FC)
ggsave('/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/endogenous_pulldown_analysis/2_plus_reps_2_plus_peps/correlation_plot/LPS vs M0 pulldown.png', width= 2, height =4)


```

## Enrichment

```{r, fig.width=2.5, fig.height=5}
# Fisher-exact GO enrichment of the crapome-filtered IP-MS pulldown gene set, run through
# the shared engine (go_enrich() in macrophage_figure_setting.R, MSigDB 2026 GO).
# process_ipms_endogenous.py exports go_gene_set.csv (dataset, gene); background = NULL
# (full human universe), matching the original "gmt" background.
# Compact manuscript panel: 6 pt fonts kept; the three ontologies stack in one
# column (facet_ncol = 1) and only the top few terms per ontology are shown.
go_gene_set <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/census-out-processing/go_gene_set.csv' |>
  read_csv(show_col_types = FALSE)

df <- go_gene_set |>
  group_by(dataset) |>
  group_modify(~ go_enrich_multi(.x$gene, onts = c("BP", "MF", "CC"))) |>
  ungroup() |>
  mutate(Term = str_wrap(Term, 30), ontology = paste(ontology, "ontology"))

df |>
  filter(adjusted_p_value < 0.05, n_proteins > 5, dataset == "Tbck") |>
  group_by(ontology) |>
  slice_min(adjusted_p_value, n = 10) |>
  ungroup() |>
  go_dotplot(
    facet = "ontology",
    facet_ncol = 1,
    size_range = c(1, 3),
    title = "Tbck pulldown GO enrichment\n(FDR < 0.05, > 5 proteins, top 5 per ontology)"
  )



ggsave(
  "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/04_IP-MS/03_analysis/Tbck_Fy2_pulldown_GO_enrichment.pdf",
  width = 2.5, height = 4.4
)


df |> write_csv('/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 5/Panels/Tbck_IP_GO-term_crapome_filter/go_term_results.csv')

```


# Multi-modal heatmap

```{r, fig.width=1.6, fig.height=4.3}
heatmap_dir <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 1_RNAseq, TMT-exp/Panels/multi_modal_heatmap/"

modalities <- c(
  "rna"          = "RNAseq",
  "qpcr"         = "qPCR",
  "western_blot" = "western blot",
  "protein"      = "whole proteome"
)
paste0(heatmap_dir, 'multi_modal_significance.csv') |> 
  read_csv(show_col_types = FALSE)  |> 
  filter(protein != "CLIC1") |> 
  mutate(
    modality = str_replace_all(modality,modalities),
    modality = factor(modality, rev(modalities)),
    condition = factor(condition, conditions)
  ) |> 
  ggplot(aes(
    x = condition,
    y = modality,
    fill = log2_fold_change,
    size = -log10(p_value)
  )) +
    geom_point(shape = 21, stroke = LINE_WIDTH) +
    facet_wrap(
      ~protein,
      ncol = 1
    ) +
    my_theme() +
    theme(
      # 4 modality rows x 7 conditions -> square cells so the dots render round.
      aspect.ratio = 4 / 7,
      panel.spacing.x = unit(0, "mm"),
      #panel.grid.major = element_line(color = "grey", linewidth = LINE_WIDTH),
      strip.placement = "outside",
      strip.text = element_text(family = FONT_FAMILY, size = FONT_SIZE, color = "black", margin = margin(b = 2)),
      legend.position = "bottom",
      legend.key.height = unit(3, "mm"),
      legend.key.width = unit(4, "mm"),
      strip.background = element_blank(),
      axis.text.x = element_text(
        color="black",angle=45, size=FONT_SIZE, family="Arial", vjust = 1, hjust=1),
      legend.box = "vertical",
      legend.box.just = "left",
      legend.box.spacing = unit(0.5, "mm"),
            plot.margin = margin(2, 2, 2, 2),
            legend.margin = margin(0, 0, 0, 0),
      legend.text = element_text(
            size=6, color = "black",
            family="Arial",
            margin = margin(l = 0)),


    ) +
    scale_fill_gradient2(
      high = "red",
      low = "blue",
      mid = "white",
      limits = c(-3, 3),
      oob = squish,
      name = expression("log"[2] * "(FC from M0)")
    ) +
    scale_size_continuous(
      range = c(1, 3),
      name = expression("-log"[10] * "(p-value)"),
      breaks = c(0, 20, 40)
    ) +
    labs(x = NULL, y = NULL)
ggsave(
  paste0(heatmap_dir, "heatmap.pdf"),
  width = 1.6,
  height = 4.3

)
```


# Multi-modal heatmap (TBCK complex) — RNA-seq + whole proteome only

```{r, fig.width=1.5, fig.height=3.5}
tbck_heatmap_dir <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 5/Panels/multi_modal_heatmap/"

# Reactivity modality labels are per-residue (e.g. "reactivity (C737)") and only exist
# for proteins with a called change, so free-y facets show each protein's own row.
# Order (top -> bottom, central dogma): RNAseq, whole proteome, reactivity.
react_data <- paste0(tbck_heatmap_dir, 'multi_modal_significance.csv') |>
  read_csv(show_col_types = FALSE) |>
  filter(condition != "TLR9") |>
  mutate(
    is_reactivity = str_detect(modality, "^reactivity"),
    modality = str_replace_all(modality, c("rna" = "RNAseq", "protein" = "whole proteome")),
    condition = factor(condition, setdiff(conditions, "TLR9")),
    protein = factor(protein, levels = c("TBCK", "CRYZL1", "C12orf4", "PPP1R21", "GATD1"))
  )
react_labels <- react_data |> filter(is_reactivity) |> pull(modality) |> unique()
react_data |>
  mutate(modality = factor(modality, levels = c(react_labels, "whole proteome", "RNAseq"))) |>
  ggplot(aes(
    x = condition,
    y = modality,
    fill = log2_fold_change
  )) +
    # RNA-seq + whole proteome: bubble size encodes significance (-log10 p).
    geom_point(
      data = \(d) filter(d, !is_reactivity),
      aes(size = -log10(p_value)),
      shape = 21, stroke = LINE_WIDTH
    ) +
    # Reactivity: no p-value; a block is drawn only where a reactivity change was
    # called. Fill (shared red/blue gradient) encodes direction/magnitude of the change.
    geom_tile(
      data = \(d) filter(d, is_reactivity),
      color = "white", linewidth = LINE_WIDTH
    ) +
    facet_wrap(
      ~protein,
      ncol = 1,
      scales = "free_y"
    ) +
    my_theme() +
    theme(
      panel.spacing.x = unit(0, "mm"),
      #panel.grid.major = element_line(color = "grey", linewidth = LINE_WIDTH),
      strip.placement = "outside",
      strip.text = element_text(family = FONT_FAMILY, size = FONT_SIZE, color = "black", margin = margin(b = 2)),
      legend.position = "bottom",
      legend.key.height = unit(3, "mm"),
      legend.key.width = unit(4, "mm"),
      strip.background = element_blank(),
      axis.text.x = element_text(
        color="black",angle=45, size=FONT_SIZE, family="Arial", vjust = 1, hjust=1),
      legend.box = "vertical",
      legend.box.just = "left",
      legend.box.spacing = unit(0.5, "mm"),
            plot.margin = margin(2, 2, 2, 2),
            legend.margin = margin(0, 0, 0, 0),
      legend.text = element_text(
            size=6, color = "black",
            family="Arial",
            margin = margin(l = 0)),


    ) +
    scale_fill_gradient2(
      high = "red",
      low = "blue",
      mid = "white",
      limits = c(-3, 3),
      oob = squish,
      name = expression("log"[2] * "(FC from M0)")
    ) +
    scale_size_continuous(
      range = c(1, 3),
      name = expression("-log"[10] * "(p-value)"),
      breaks = c(0, 1, 2, 3, 4, 5)
    ) +
    labs(x = NULL, y = NULL)
ggsave(
  paste0(tbck_heatmap_dir, "heatmap.pdf"),
  width = 1.5,
  height = 3.5

)
```




