

```{r}

source('../macrophage_figure_setting.R')
library(colorspace)
library(ggalluvial)
results_dir <- "/Users/henrysanford/dev/test_data/macrophage/02_reactivity/03_rc_analysis/20250225_3reps/06_results"


protein_list_dir <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Exhaustion manuscript/02_Mass-spectrometry data/Reference lists/"
```

## Cysteines quantified barplot

```{r}
reactivity_vs_wp_dir <- paste0(results_dir, "/03_reactivity_vs_wp")
#paste0(reactivity_vs_wp_dir, "/reactivity_changes_long_format.csv") %>% 
'/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/02_reactivity/03_rc_analysis/reactivity_output/test/median_for_5+_peptides/output/reactivity_changes_long_format.csv' %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(
    LFC_tmt_abpp = log2( rc_percent_control / 100),
    LFC_WP = log2( wp_percent_control / 100),
    direction_of_reactivity_change = str_to_title(direction_of_reactivity_change)
    
    )-> full_df
full_df
```



```{r, fig.height=1.8, fig.width=2}
AXIS_TXT_MRGN = -5


quant_colors <- c(
  "reactivity changes" = "indianred2",
  "analyzed" = "grey39",
  "not analyzed" = "grey90"
)

full_df %>%
  group_by(protein, rc_n_peptides) %>%
  summarise(rc_any = any(reactivity_change)) %>%
  mutate(
    reactivity_change = case_when(
      rc_n_peptides == 1 ~ "not analyzed",
      rc_any == TRUE ~ "reactivity changes",
      TRUE ~ "analyzed"
    ),
    binned_quantified_peptides = case_when(
      (rc_n_peptides > 6) ~ "7+",
      (rc_n_peptides > 4) ~ "5-6",
      TRUE ~ as.character(rc_n_peptides)
    ),
    reactivity_change = factor(reactivity_change, levels = rev(names(quant_colors)))
  ) %>%
  arrange(reactivity_change, ascending = TRUE) %>%
  distinct(.keep_all = TRUE) %>%
  ggplot(aes(x = binned_quantified_peptides, fill = reactivity_change)) +
  geom_bar(show.legend = FALSE, width = BARPLOT_WIDTH) +
  scale_fill_manual(values = quant_colors,
                    breaks = names(quant_colors),
                    name = "") +
  guides(fill = guide_legend(ncol = 1, hjust = 1)) +
  scale_y_break(c(200, 200), scales = 1 / 1.2, expand = c(0, 0)) +
  scale_x_discrete() +
  my_theme() +
  theme(
    legend.position =  "top",
    legend.justification = "left",
    legend.key.spacing = unit(-0.1, "mm"),
    legend.key.spacing.x = unit(500, "mm"),
    legend.margin = margin(b = -5),
    # reverse annoying ggbreak changes
    axis.line.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title.x = element_text(
      margin = margin(t = AXIS_TXT_MRGN),
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE,
      hjust = 0.5
    ),
    axis.title.y = element_text(
      margin = margin(r = AXIS_TXT_MRGN),
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE,
      vjust = 0.5
    ),
    panel.border = element_blank(),
    axis.line.x = AXIS_LINE,
    axis.line.y = AXIS_LINE
  ) +
  ylab("number of proteins") +
  xlab("quantified cysteines") +
  ylim(0, 2400)

save_plot(
  paste0(results_dir, "/07_barplots/num_peptides_rc"),
  height = 2,
  width = 2
)
```


## WP RC scatterplot

### All cysteines

```{r}
reactivity_vs_wp_dir <- paste0(results_dir, "/03_reactivity_vs_wp")

paste0(
  reactivity_vs_wp_dir,
  "/all-conditions-vs_M0_rc_wp_comparison.csv"
) %>%
  read_csv() %>%
  drop_na("LFC_WP") %>%
  filter(LFC_tmt_abpp != Inf) -> df

max(df$LFC_tmt_abpp) %>% print()
max(df$LFC_WP) %>% print()
```


```{r, fig.width=4, fig.height=2}
  
  scatterplot_palette <- c("Higher" = SIG_UP_COL, 
                  "Lower" = SIG_DOWN_COL, 
                  "Unchanged or\nexpression based" = UNCHANGED_COL)
  
  map <- setNames(c("Higher",
                    "Lower",
                    "Unchanged or\nexpression based"),
                  c("Higher",
                    "Lower", 
                    "Unchanged or expression based"))
  
protein_scatterplot <- function(
    full_df,
    gene_name, 
    fil_condition, 
    dir
    ){
  dir <- paste0(
    dir, 
    "/",
    fil_condition,
    "_rc_vs_wp"
  )
  
  full_df$cysteine_reactivity_regulation <- map[unlist(
    full_df$cysteine_reactivity_regulation
    )]
  
  full_df %>% 
    filter(condition == fil_condition) %>% 
    mutate(name = paste0(protein, "_", residue)) -> df
  
  # the limit of the plot should be the maximum
  # whole proteome or reactivity LFC
  plot_lim <- max(
    max(abs(df$LFC_WP)),
    max(abs(df$LFC_tmt_abpp))
  )
  breaks <- c(-round(plot_lim):round(plot_lim))
  
  df %>% 
    filter(protein == gene_name) -> label_data
  high_cutoff = log2(1.6)
  low_cutoff = log2(1/1.6)
  df %>% 
    ggplot(aes(x=LFC_tmt_abpp, 
               y=LFC_WP, 
               fill=cysteine_reactivity_regulation)) + 
    geom_point(
               color="black" ,
               size = 1,
               stroke=POINT_STROKE / 2,
               shape=21, 
               alpha=0.6,
               show.legend = FALSE) + 
    geom_point(data=label_data, 
               color="black" ,
               size = 1,
               stroke=POINT_STROKE / 2,
               shape=21, 
               show.legend = FALSE) + 
    geom_rect(data = df[1,],
              aes(xmin = -Inf, 
                  xmax = -1, 
                  ymax = Inf, 
                  ymin = low_cutoff), 
              color = "black", 
              linewidth = LINE_WIDTH,
              fill="transparent", 
              linetype = "dotted") +
    geom_rect(data = df[1,],aes(xmin = 1, 
                                xmax = Inf, 
                                ymax = high_cutoff, 
                                ymin = -Inf),
              color = "black", 
              linewidth = LINE_WIDTH,
              fill="transparent", 
              linetype = "dotted") +
    scale_fill_manual(values = scatterplot_palette) +
    scale_x_continuous(breaks = breaks,
                       limits = c(-plot_lim, plot_lim)) +
    scale_y_continuous(breaks = breaks,
                       limits = c(-plot_lim, plot_lim)) +
    guides(fill=guide_legend(
      title=paste0("Cysteine reactivity in ", fil_condition))) + 
   geom_text_repel(data = label_data, 
                     aes(label=name), 
                    color = "black",
                     size=1.76,
                    segment.size = LINE_WIDTH,
                     min.segment.length = unit(0, 'lines'), 
                     nudge_y=2.5, 
                     show.legend = FALSE) + 
    xlab(paste0("cysteine reactivity log2(", fil_condition,"/", control_condition ,")")) +
    ylab(paste0("protein expression log2(", fil_condition,"/", control_condition ,")")) +
    my_theme() +
    theme(aspect.ratio=1, 
          legend.key.spacing.y = unit(-2,"mm"),
          legend.key = element_blank(),)-> plot
    print(plot)
    save_plot(dir, 2, 2)
}
for(condition in conditions){
 protein_scatterplot(
  df,
  c(""),
  condition,
  reactivity_vs_wp_dir
  ) 
}
```



## Reactivity dotplots 






### All conditions


```{r, fig.height=2.3, fig.width=4.1}
lbl_nudge <- 0.4
reactivity_vs_wp_dir <- paste0(results_dir, "/03_reactivity_vs_wp")
fil_condition <- "D8C"
rc_dotplot <- function(df,
                       targets,
                       title,
                       fil_condition,
                       control_condition,
                       output_dir,
                       ratio_aspect = 1,
                       ht = 4,
                       wd = 2,
                       y_max = 3,
                       y_min = -3) {
  df %>%
    filter(condition == fil_condition) %>%
    filter(protein %in% targets) %>%
    mutate(
      `Cysteine reactivity in D8C` = case_when(
        reactivity_change != TRUE ~ "Unchanged",
        LFC_tmt_abpp > 0 ~ "Higher",
        LFC_tmt_abpp < 0 ~ "Lower",
        TRUE ~ "Unchanged"
      ),
      protein = factor(protein, levels = targets)
    ) -> plt_df
  
  lbl_data <- plt_df %>%
    filter(reactivity_change)
  
  wp_data <- plt_df %>%
    select(uniprot, protein, condition, LFC_WP) %>%
    mutate(`Cysteine reactivity in D8C` = "Protein expression") %>%
    distinct()
  
  plt_df %>%
    arrange(by = "protein") %>%
    ggplot(aes(x = protein, y = LFC_tmt_abpp, fill = `Cysteine reactivity in D8C`)) +
    geom_hline(yintercept = c(1, -1),
               linetype = "dashed",
               size = LINE_WIDTH) +
    # cysteine data
    geom_point(
      size = POINT_SIZE - 0.75,
      shape = 21,
      color = "black",
      alpha = 0.9,
      stroke = POINT_STROKE
    ) +
    # residue labels
    geom_text(
      data = lbl_data,
      mapping = aes(
        y = LFC_tmt_abpp + case_when(
          LFC_tmt_abpp > 0 ~ lbl_nudge,
          LFC_tmt_abpp < 0 ~ -lbl_nudge,
          TRUE ~ lbl_nudge
        ),
        color = `Cysteine reactivity in D8C`,
        label = residue
      ),
      size = FONT_SIZE,
      size.unit = "pt",
      show.legend = FALSE
    ) +
    # whole proteome data (triangle)
    geom_point(
      data = wp_data,
      aes(x = protein, y = LFC_WP),
      shape = 24,
      color = "black",
      size = POINT_SIZE - 0.75,
      stroke = POINT_STROKE,
      alpha = 0.9
    ) +
    # aesthetics
    my_theme() +
    scale_fill_manual(values = regulation_colors,
                      breaks = names(regulation_colors)) +
    scale_color_manual(values = regulation_colors,
                       breaks = names(regulation_colors)) +
    scale_y_continuous(breaks = round(y_min):round(y_max),
                       limits = c(y_min, y_max)) +
    guides(fill = guide_legend(byrow = TRUE,
                               nrow = TRUE)) +
    theme(
      axis.text.x = element_text(
        color = "black",
        angle = 45,
        size = FONT_SIZE,
        family = "Arial",
        vjust = 1,
        hjust = 1
      ),
      legend.key.spacing.y = unit(-2.5, "mm"),
      legend.margin = margin(b = -15),
      legend.position = "top",
      #aspect.ratio = ratio_aspect,
      legend.title = element_text(margin = margin(b = -2)),
      legend.key = element_blank(),
      legend.title.position = "top",
      legend.key.spacing = unit(0, "mm"),
      legend.spacing.y = unit(-0, "mm"),
      legend.spacing = unit(0, "mm")
    ) +
    labs(y = substitute(
      paste('cysteine reactivity, log'[2],
            "(", n, "/", m, ")"),
      list(n = fil_condition, m = control_condition)
    ),
    x = "") +
    guides(fill = guide_legend(ncol = 2)) -> plot
  print(plot)
  save_plot(paste0(output_dir, title),
            height = ht,
            width = wd)
}




```

```{r, fig.height=2.3, fig.width=4}
trafficking_proteins <- c("DCTN1","AP3D1","VPS13C","SNX6","TMF1","GOLGA3","GOLGB1","STXBP3","TRIP11","NBAS","VPS11","VPS18","PLEKHM2","STAB1","DIAPH2","CORO1C")

rc_dotplot(
  df = full_df,
  targets = trafficking_proteins,
  title = "Trafficking proteins",
  fil_condition = "TLR4",
  control_condition = "M0",
  output_dir = paste0(results_dir, "/06_dotplots/"),
  #y_max = 10,
  #y_min = -5,
  y_max = 4.3,
  y_min = -2.2,
  ht = 2.3,
  wd = 4
)

full_df %>%
  filter(
    condition =="TLR4"
  ) 
```

```{r, fig.width=1, fig.height=2.8}

macrophage_label <- "Macrophage, TLR4/M0"
tcell_label <- "T cells, Activated/Control"
# data from cell 2020
data <- data.frame(
  uniprot = "P00390",
  residue = c("C102","C377","C107","C484","C328","C278"),
  protein = "T cells, Activated/Control",
  condition = "TLR4",
  reactivity_change = c(TRUE, FALSE, FALSE, FALSE, FALSE, FALSE ),
  direction_of_reactivity_change = c("Higher", "unchanged", "unchanged", "unchanged", "unchanged", "unchanged"),
  LFC_tmt_abpp = log2(c(4.29,0.84,0.68,0.54,0.54,0.47)),
  LFC_WP = log2(0.59)
)
# combine with macrophage data
full_df |> 
  select(uniprot,residue, protein, condition, reactivity_change,direction_of_reactivity_change ,LFC_tmt_abpp,LFC_WP) |> 
  filter(protein == "GSR", condition == "TLR4") |> 
  mutate(protein = "Macrophage, TLR4/M0") |> 
  rbind(data) |> 
  rc_dotplot(
  targets = c(macrophage_label, tcell_label),
  title = "GSR",
  fil_condition = "TLR4",
  control_condition = "M0",
  output_dir = paste0(results_dir, "/06_dotplots/"),
  #y_max = 10,
  #y_min = -5,
  y_max = 2.6,
  y_min = -1.5,
  ht = 2.8,
  wd = 1
)

results_dir
  
  
```





```{r, fig.height=1.5, fig.width=2}
GAP_GEF_proteins <- c(
  #GAPs
"ARAP1" , "TBC1D15", "ARHGAP4",
# GEFs
"TIAM1", "TRIO", "GAPVD1"
)


lbl_nudge <- 0.4

fil_condition <- "D8C"
rc_dotplot_t <- function(df,
                       targets,
                       title,
                       fil_condition,
                       control_condition,
                       output_dir,
                       ratio_aspect = 1,
                       ht = 4,
                       wd = 2,
                       y_max = 3,
                       y_min = -3) {
  df %>%
    filter(condition == fil_condition) %>%
    filter(protein %in% targets) %>%
    mutate(
      `Cysteine reactivity in D8C` = case_when(
        reactivity_change !=TRUE ~ "Unchanged",
        LFC_tmt_abpp > 0 ~ "Higher",
        LFC_tmt_abpp < 0 ~ "Lower",
        TRUE ~ "Unchanged"
      ),
      protein = factor(protein, levels = rev(targets))
    ) -> plt_df
  
  lbl_data <- plt_df %>%
    filter(reactivity_change)
  
  wp_data <- plt_df %>%
    select(uniprot, protein, condition, LFC_WP) %>%
    mutate(`Cysteine reactivity in D8C` = "Protein expression") %>%
    distinct()
  
  plt_df %>%
    arrange(by = "protein") %>%
    ggplot(aes(y = protein, x = LFC_tmt_abpp, fill = `Cysteine reactivity in D8C`)) +
    geom_vline(xintercept = c(1, -1),
               linetype = "dashed",
               size = LINE_WIDTH) +
    # cysteine data
    geom_point(
      size = POINT_SIZE - 0.75,
      shape = 21,
      color = "black",
      alpha = 0.9,
      stroke = POINT_STROKE
    ) +
    # residue labels
    geom_text(
      data = lbl_data,
      mapping = aes(
        x = LFC_tmt_abpp + case_when(
          LFC_tmt_abpp > 0 ~ lbl_nudge,
          LFC_tmt_abpp < 0 ~ -lbl_nudge,
          TRUE ~ lbl_nudge
        ),
        color = `Cysteine reactivity in D8C`,
        label = residue
      ),
      size = FONT_SIZE,
      size.unit = "pt",
      show.legend = FALSE
    ) +
    # whole proteome data (triangle)
    geom_point(
      data = wp_data,
      aes(y = protein, x = LFC_WP),
      shape = 24,
      color = "black",
      size = POINT_SIZE - 0.75,
      stroke = POINT_STROKE,
      alpha = 0.9
    ) +
    # aesthetics
    my_theme() +
    scale_fill_manual(values = regulation_colors,
                      breaks = names(regulation_colors)) +
    scale_color_manual(values = regulation_colors,
                       breaks = names(regulation_colors)) +
    scale_x_continuous(breaks = round(y_min):round(y_max),
                       limits = c(y_min, y_max)) +
    guides(fill = guide_legend(byrow = TRUE,
                               nrow = TRUE)) +
     theme(
    #   axis.text.x = element_text(
    #     color = "black",
    #     angle = 45,
    #     size = FONT_SIZE,
    #     family = "Arial",
    #     vjust = 1,
    #     hjust = 1
    #   ),
      legend.key.spacing.y = unit(-2.5, "mm"),
      legend.margin = margin(b = -15),
      legend.position = "top",
      #aspect.ratio = ratio_aspect,
      legend.title = element_text(margin = margin(b = -2)),
      legend.key = element_blank(),
      legend.title.position = "top",
      legend.key.spacing = unit(0, "mm"),
      legend.spacing.y = unit(-0, "mm"),
      legend.spacing = unit(0, "mm")
    ) +
    labs(x = substitute(
      paste('cysteine reactivity, log'[2],
            "(", n, "/", m, ")"),
      list(n = fil_condition, m = control_condition)
    ),
    y = NULL) +
    guides(fill = guide_legend(ncol = 2)) -> plot
  print(plot)
  save_plot(paste0(output_dir, title),
            height = ht,
            width = wd)
}

rc_dotplot_t(
  df = full_df,
  targets = GAP_GEF_proteins,
  title = "GAP_GEF_proteins_t",
  fil_condition = "TLR4",
  control_condition = "M0",
  output_dir = paste0(results_dir, "/06_dotplots/"),
  y_max = 5.3,
  y_min = -2.2,
  ht = 1.7,
  wd = 2.3
)
```




```{r, fig.width = 5, fig.height = 2.349}
#TODO add full_df as argument
facetted_multi_condition_dotplot <-
  function(targets,
           display_conditions,
           title,
           additional_cysteines_to_label = c(),
           fig_height,
           fig_width,
           output_dir,
           control_condition = "M0",
           y_min = -4.5,
           y_max = 2) {
    
    max_val <- max(full_df$LFC_tmt_abpp[is.finite(full_df$LFC_tmt_abpp)], na.rm = TRUE)

  full_df <- full_df |>
    mutate(LFC_tmt_abpp = case_when(
      LFC_tmt_abpp == Inf  ~  max_val,
      LFC_tmt_abpp == -Inf ~ -max_val,
      .default = LFC_tmt_abpp
    ))
  
  full_df |> filter(protein == "ARHGAP31") |> arrange(desc(LFC_tmt_abpp)) |> print()
    full_df %>%
      filter(protein %in% targets,
             condition %in% display_conditions) %>%
      mutate(`PPAT reactivity in D8C` = direction_of_reactivity_change,
      condition = factor(condition, levels = names(cols)),
      protein = factor(protein, levels = targets)) -> plt_df
    
    plt_df %>%
      filter(reactivity_change == TRUE |
               residue %in% additional_cysteines_to_label) -> lbl_data
    
    plt_df %>% select(uniprot, protein, condition, LFC_WP) %>% 
      mutate(`PPAT reactivity in D8C` = "Protein expression") %>%
      distinct() -> wp_data
    
    plt_df$protein <- factor(plt_df$protein, levels = targets)
    plt_df %>%
      ggplot(aes(x = condition, y = LFC_tmt_abpp, fill = `PPAT reactivity in D8C`)) +
      geom_hline(yintercept =c(-1, 1),
                 linetype = "dashed",
                 size = LINE_WIDTH) +
      # cysteine data
      geom_point(
        size = POINT_SIZE - 0.75,
        shape = 21,
        color = "black",
        alpha = 0.9,
        stroke = POINT_STROKE
      ) +
      # residue labels
      geom_text(
        data = lbl_data,
        mapping = aes(
          y = LFC_tmt_abpp + case_when(LFC_tmt_abpp > 0 ~ 0.3,
                                       LFC_tmt_abpp < 0 ~ -0.3,
                                       TRUE ~ 0.3),
          color = `PPAT reactivity in D8C`,
          label = residue
        ),
        size = FONT_SIZE,
        size.unit = "pt",
        show.legend = FALSE
      ) +
      # whole proteome data (triangle)
      geom_point(
        data = wp_data,
        aes(x = condition, y = LFC_WP),
        shape = 24,
        color = "black",
        size = POINT_SIZE - 0.75,
        stroke = POINT_STROKE,
        alpha = 0.9
      ) +
      # aesthetics
      my_theme() +
      scale_fill_manual(values = regulation_colors,
                        breaks = names(regulation_colors)) +
      scale_color_manual(values = regulation_colors,
                         breaks = names(regulation_colors)) +
      scale_y_continuous(breaks = round(y_min):round(y_max),
                         limits = c(y_min, y_max)) +
      theme(
        legend.key.spacing.y = unit(-2.5, "mm"),
        legend.margin = margin(b = -15),
        legend.position = "top",
        strip.text = element_text(
          family = FONT_FAMILY,
          color = "black",
          size = FONT_SIZE,
          margin = margin(t = 0)
        ),
        strip.background = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(0, "mm"),
        legend.title = element_text(margin = margin(b = -2)),
        legend.key = element_blank(),
        legend.title.position = "top",
        legend.key.spacing = unit(0, "mm"),
        legend.spacing.y = unit(-0, "mm"),
        legend.spacing = unit(0, "mm")
      ) +
      labs(y = substitute(
        paste('cysteine reactivity, log'[2], "(FC from ", m,")"),
        list(n = fil_condition, m = control_condition)
      ),
      x = NULL,
      fill = NULL) +
      facet_wrap( ~ protein, ncol = length(unique(plt_df$protein)), strip.position = "bottom") +
      guides(fill = guide_legend(ncol = 2)) +
    theme(
      axis.text.x = element_text(
        color = "black",
        angle = 45,
        size = FONT_SIZE,
        family = "Arial",
        vjust = 1,
        hjust = 1
      ) ,     
      panel.border = element_rect(color = "black", linetype = "solid", linewidth = LINE_WIDTH))-> plot
    print(plot)
    save_plot(
      paste0(output_dir, title),
      height = fig_height,
      width = fig_width
    )
  }

autophagy_proteins <- c(
  "AKAP11","ATG7","ATG9A","EPG5","LRBA","PIK3C2A","VPS13C","UVRAG"
)
full_df <- full_df %>%
  mutate(direction_of_reactivity_change = case_when(
    direction_of_reactivity_change == "unchanged" ~ "Unchanged",
    TRUE ~ direction_of_reactivity_change
  ))
```


```{r, fig.width = 4.2, fig.height = 2.349}

facetted_multi_condition_dotplot(
  targets = c("ATG7","ATG9A","EPG5","FYCO1","NRBF2","PIK3C2A","UVRAG","VPS13C"),
  display_conditions = c("TLR4", "TLR8","STING"),
  additional_cysteines_to_label = c(),
  title = "trim_48STING",
  output_dir = paste0(results_dir, "/06_dotplots/"),
  fig_height = 2.349, 
  fig_width =4.2,
  y_max = 4.2,
  y_min = -1
)


```

### CDC 42 interactors

```{r, fig.width=6, fig.height=3.8}

facetted_multi_condition_dotplot <-
  function(targets,
           display_conditions,
           title,
           additional_cysteines_to_label = c(),
           fig_height,
           fig_width,
           output_dir,
           control_condition = "M0",
           y_min = -4.5,
           y_max = 2) {
    
    max_val <- max(full_df$LFC_tmt_abpp[is.finite(full_df$LFC_tmt_abpp)], na.rm = TRUE)

  full_df <- full_df |>
    mutate(LFC_tmt_abpp = case_when(
      LFC_tmt_abpp == Inf  ~  max_val,
      LFC_tmt_abpp == -Inf ~ -max_val,
      .default = LFC_tmt_abpp
    ))
  
  full_df |> filter(protein == "ARHGAP31") |> arrange(desc(LFC_tmt_abpp)) |> print()
    full_df %>%
      filter(protein %in% targets,
             condition %in% display_conditions) %>%
      mutate(`PPAT reactivity in D8C` = direction_of_reactivity_change,
      condition = factor(condition, levels = names(cols)),
      protein = factor(protein, levels = targets)) -> plt_df
    
    plt_df %>%
      filter(reactivity_change == TRUE |
               residue %in% additional_cysteines_to_label) -> lbl_data
    
    plt_df %>% select(uniprot, protein, condition, LFC_WP) %>% 
      mutate(`PPAT reactivity in D8C` = "Protein expression") %>%
      distinct() -> wp_data
    
    plt_df$protein <- factor(plt_df$protein, levels = targets)
    plt_df %>%
      ggplot(aes(x = condition, y = LFC_tmt_abpp, fill = `PPAT reactivity in D8C`)) +
      geom_hline(yintercept =c(-1, 1),
                 linetype = "dashed",
                 size = LINE_WIDTH) +
      # cysteine data
      geom_point(
        size = POINT_SIZE - 0.75,
        shape = 21,
        color = "black",
        alpha = 0.9,
        stroke = POINT_STROKE
      ) +
      # residue labels
      geom_text(
        data = lbl_data,
        mapping = aes(
          y = LFC_tmt_abpp + case_when(LFC_tmt_abpp > 0 ~ 0.3,
                                       LFC_tmt_abpp < 0 ~ -0.3,
                                       TRUE ~ 0.3),
          color = `PPAT reactivity in D8C`,
          label = residue
        ),
        size = FONT_SIZE,
        size.unit = "pt",
        show.legend = FALSE
      ) +
      # whole proteome data (triangle)
      geom_point(
        data = wp_data,
        aes(x = condition, y = LFC_WP),
        shape = 24,
        color = "black",
        size = POINT_SIZE - 0.75,
        stroke = POINT_STROKE,
        alpha = 0.9
      ) +
      # aesthetics
      my_theme() +
      scale_fill_manual(values = regulation_colors,
                        breaks = names(regulation_colors)) +
      scale_color_manual(values = regulation_colors,
                         breaks = names(regulation_colors)) +
      scale_y_continuous(breaks = round(y_min):round(y_max),
                         limits = c(y_min, y_max)) +
      theme(
        legend.key.spacing.y = unit(-2.5, "mm"),
        legend.margin = margin(b = -15),
        legend.position = "top",
        strip.text = element_text(
          family = FONT_FAMILY,
          color = "black",
          size = FONT_SIZE,
          margin = margin(t = 0)
        ),
        strip.background = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(0, "mm"),
        legend.title = element_text(margin = margin(b = -2)),
        legend.key = element_blank(),
        legend.title.position = "top",
        legend.key.spacing = unit(0, "mm"),
        legend.spacing.y = unit(-0, "mm"),
        legend.spacing = unit(0, "mm")
      ) +
      labs(y = substitute(
        paste('cysteine reactivity, log'[2], "(FC from ", m,")"),
        list(n = fil_condition, m = control_condition)
      ),
      x = NULL,
      fill = NULL) +
      facet_wrap( ~ protein, ncol = length(unique(plt_df$protein)), strip.position = "bottom") +
      guides(fill = guide_legend(ncol = 2)) +
    theme(
      axis.text.x = element_text(
        color = "black",
        angle = 45,
        size = FONT_SIZE,
        family = "Arial",
        vjust = 1,
        hjust = 1
      ) ,     
      panel.border = element_rect(color = "black", linetype = "solid", linewidth = LINE_WIDTH))-> plot
    print(plot)
    save_plot(
      paste0(output_dir, title),
      height = fig_height,
      width = fig_width
    )
  }

fil_condition <- "TLR4"
facetted_multi_condition_dotplot(
  targets = c(
    "ARHGAP21",
    "ARHGAP31",
    "ARHGAP4",
    "ARHGEF7",
    "DCTN1",
    "DIAPH2",
    "DNMBP",
    "DOCK4",
    "DOCK7",
    "DOCK8",
    "IQGAP1",
    "PREX1",
    "TIAM1",
    "TNK2"
  ),
  display_conditions = c("TLR1-2", "TLR4"),
  additional_cysteines_to_label = c(),
  title = "cdc42_interactors",
  output_dir = paste0(results_dir, "/06_dotplots/"),
  fig_height = 3.8,
  fig_width = 6,
  y_max = 7.5,
  y_min = -1.1
)
full_df
results_dir
```





```{r}
library(tidyverse)
reactivity_changes <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/02_reactivity/03_rc_analysis/reactivity_output/20251023_ratio_directionality/median_for_5+_peptides/output/reactivity_changes_long_format.csv' |> 
  read_csv(show_col_types = FALSE)

reactivity_changes |> 
  filter(protein =="ARHGAP31") |> 
  arrange(desc(rc_percent_control))
```







```{r, fig.width = 5, fig.height = 2.349}
phos_homeostasis_proteins <- c("INPP5D","FIG4","PIK3C2A","PI4KA","PLCB2","PLCB3","VAC14","PIP5K1A")
facetted_multi_condition_dotplot(
  targets = phos_homeostasis_proteins,
  display_conditions = c("TLR4", "TLR7", "STING"),
  additional_cysteines_to_label = c(),
  title = "phos_homeostasis_proteins_TLR4_TLR7_STING",
  output_dir = paste0(results_dir, "/06_dotplots/"),
  fig_height = 2.349, 
  fig_width = 5,
  y_max =4.2,
  y_min = -1.9
)

```


```{r, fig.width = 4.5, fig.height = 2.349}
facetted_multi_condition_dotplot(
  targets = autophagy_proteins,
  display_conditions = c("TLR4", "TLR7", "STING"),
  additional_cysteines_to_label = c(),
  title = "autophagy proteins_TLR4_TLR7_STING",
  output_dir = paste0(results_dir, "/06_dotplots/"),
  fig_height = 2.349, 
  fig_width =4.5,
  y_max = 4.2,
  y_min = -1
)
```






```{r, fig.height=2.4, fig.width=2.7}
multi_condition_rc_plot <- function(df,
                                    proteins,
                                    output_dir,
                                    ht = 2.5,
                                    wd = 2,
                                    rotate_x_labels = TRUE,
                                    control_condition = "M0",
                                    brk_size = 100) {
df %>%
    filter(protein %in% proteins) %>%
    mutate(residue_loc = 
             str_split_fixed(
               str_split_fixed(residue, "C", 2)[, 2], ";", 2)[, 1] %>% 
             as.numeric()) -> map2k_df
  
  map2k_df %>% filter(reactivity_change == "True") -> label_data
  
  max_tmt <- max(map2k_df$LFC_tmt_abpp)
  map2k_df %>%
    filter(reactivity_change != "True") %>%
    ggplot(aes(x = reorder(residue, residue_loc), y = LFC_tmt_abpp)) +
    theme_minimal() +
    geom_hline(
      yintercept = 0,
      linetype = "solid",
      size = LINE_WIDTH,
      color = "black"
    ) +
    labs(
      x = NULL,
      y = substitute(
      paste('cysteine reactivity, log'[2],
            "(FC from ", m, ")"),
      list(m = control_condition)
    )
    ) +
    geom_beeswarm(
      size = POINT_SIZE - 1,
      shape = 21,
      col = "black",
      show.legend = FALSE,
      fill = "lightgrey",
      width = 0.5,
      height = 0,
      stroke = POINT_STROKE,
      alpha = 0.7
    ) +
    # geom_text(
    #   data = label_data,
    #   mapping = aes(
    #     x = residue,
    #     y = LFC_tmt_abpp + (1.25 / 16) * max_tmt,
    #     label = condition,
    #     color = condition
    #   ),
    #   inherit.aes = FALSE,
    #   show.legend = FALSE,
    #   size = FONT_SIZE,
    #   size.unit = "pt"
    # ) +
    geom_beeswarm(
      data = label_data,
      mapping = aes(fill = condition),
      size = POINT_SIZE - 1,
      shape = 21,
      col = "black",
      alpha = 0.7,
      corral.width = 2,
      cex = 3,
      show.legend = FALSE,
      stroke = POINT_STROKE / 2
    ) +
    scale_fill_manual(values = cols) +
    scale_color_manual(values = cols) +
    my_theme() +
    theme(
      strip.background = element_blank(),
      panel.spacing = unit(2, "mm"),
      strip.text = element_text(
        color = "black",
        size = 6,
        family = "Arial"
      ),
      panel.grid.major = element_line(colour = "grey90", linewidth=LINE_WIDTH),
        panel.border =element_blank(),
      axis.line.x = element_line(color = "black", linetype = "solid", linewidth = LINE_WIDTH),
      axis.line.y = element_line(color = "black", linetype = "solid", linewidth = LINE_WIDTH)
    ) +
    scale_y_continuous(breaks = seq.int(0, 4000, by = brk_size)) +
    facet_grid(cols = vars(protein),space = "free", scales = "free") -> plt
  if (rotate_x_labels) {
    plt <- plt +
      theme(
        axis.text.x = element_text(
          color = "black",
          angle = 45,
          size = 5,
          family = "Arial",
          vjust = 1,
          hjust = 1
        )
      )
  }
  
  print(plt)
  save_plot(paste0(output_dir, paste(proteins, collapse = "_")),
            height = ht,
            width = wd)
}

df <-paste0(reactivity_vs_wp_dir,
         "/rc_df_long_format.csv") %>%
    read.csv() %>% filter(condition != "TLR9")

multi_condition_rc_plot(
  df = df,
  proteins = autophagy_proteins,
  output_dir = paste0(results_dir, "/06_dotplots/"),
  ht = 2.1, wd = 2.7,
  rotate_x_labels = TRUE,
  brk_size = 1
)





```
## boxplot

```{r, fig.height=2, fig.width=6}
expression_and_reactivity_boxplot_all_conditions <-
  function(condition_to_display,
           control_condition,
           proteins_to_plot,
           whole_proteome_table,
           reactivity_table,
           ht,
           wd) {
    cysteine_seperator <- ", " # for aggregated cysteines
    
    si_matrix <- whole_proteome_table %>%
      # select columns needed for plot
      select(starts_with(c(
        conditions, c("protein", control_condition)
      ))) %>%
      filter(protein %in% proteins_to_plot) %>%
      column_to_rownames("protein")
    
    # calculate fold change to median of control condition
    wp_df <- (si_matrix / apply(si_matrix[, grepl(control_condition, colnames(si_matrix))], 1, median, na.rm =TRUE)) %>%
      rownames_to_column("protein") %>%
      # pivot to long format for plotting
      pivot_longer(cols = -protein,
                   names_to = "condition_replicate",
                   values_to = "channel_ratio") %>%
      # drop replicates where the protein was not quantified
      drop_na() %>%
      separate_wider_delim(
        cols =condition_replicate,
        delim = "_",
        names = c("condition", "technical_replicate", "experiment"),
        too_many = "drop"
      ) %>% mutate(
        residue = "C0",
        regulation = "Protein expression",
        reactivity_change = FALSE
      ) 
    
    si_matrix <- reactivity_table %>%
      filter(protein %in% proteins_to_plot) %>%
      # select columns needed for plot
      select(starts_with(c(
        conditions, c("protein", "residue", control_condition)
      ))) %>% mutate(residue = gsub(';', cysteine_seperator ,residue)) %>%
      # concat protein and residue
      unite(rowname, protein, residue) %>%
      column_to_rownames()
    # calculate fold change to median of control condition
    df <- (si_matrix / apply(si_matrix[, grepl(control_condition, colnames(si_matrix))], 1, median, na.rm =TRUE)) %>%
      rownames_to_column("protein_residue") %>%
      # pivot to long format for plotting
      pivot_longer(cols = -protein_residue,
                   names_to = "condition_replicate",
                   values_to = "channel_ratio") %>%
      # drop missing values
      drop_na() %>%
      separate_wider_delim(
        cols = c("protein_residue"),
        delim = "_",
        names = c("protein", "residue")
      ) %>%
      separate_wider_delim(
        cols = c("condition_replicate"),
        delim = "_",
        names = c("condition", "experiment", "technical_replicate"),
        too_many = "drop"
      ) %>% merge(
      y = full_df %>%
        select(protein, residue, reactivity_change, condition) %>%
        mutate(residue =gsub(",", cysteine_seperator, residue)),
      by = c("protein", "residue", "condition"),
      all.x = TRUE
    ) %>% mutate(regulation = "Reactivity") %>%
      rbind(wp_df) %>%
      mutate(
        # converting strings to factors so proteins and conditions have correct order
        protein = factor(protein, levels = proteins_to_plot),
        condition = factor(condition, levels = c(control_condition, conditions)),
        # get numeric index of cysteine for sorting
        residue_loc = str_split_fixed(residue, "C", 2)[, 2],
        residue_loc = str_split_fixed(residue_loc, cysteine_seperator, 2)[, 1] %>% as.numeric(),
        # format reactivity change boolean to string
        reactivity_change = as.character(reactivity_change), 
        reactivity_change = replace_na(reactivity_change, "FALSE")
      ) %>%
      filter(condition != "TLR9")
    
    # set plot dimensions
    # adding extra buffer on the bottom for annotation
    max_plot_value <- max(log2(df$channel_ratio)) * 1.1
    min_plot_value <- min(log2(df$channel_ratio)) * 1.2
    margin_value <- min_plot_value  + 0.05 * (max_plot_value - min_plot_value)
    
    plot <- df %>%
      unique() %>%
      ggplot(aes(
        x = reorder(residue, residue_loc),
        fill = condition,
        y = log2(channel_ratio),
        shape = regulation
      )) +
      geom_boxplot(
        outliers = FALSE,
        width = 0.7,
        lwd = LINE_WIDTH,
        show.legend = FALSE,
        alpha = 0.6,
        position = position_dodge(width = 0.8)
      ) +
      # replicate level LFC values
      geom_beeswarm(
        col = "black",
        stroke = POINT_STROKE,
        show.legend = FALSE,
        size = 0.9,
        alpha = 0.7,
        dodge.width = 0.8
      ) +
      facet_rep_grid(~ protein, scales = "free_x", space = "free", ) +
      my_theme() +
      theme(
        plot.title = element_text(
          size = 6,
          color = "black",
          family = "Helvetica",
          hjust = 0.5
        ),
        strip.background = element_blank(),
        panel.spacing = unit(2, "mm"),
        strip.text = element_text(
          size = 8,
          family = "Helvetica",
          color = "black"
        ),
        strip.clip = "off",
        panel.grid.major.y = element_line(linewidth = LINE_WIDTH, linetype = "solid", color = "grey90")
      ) +
      # boxes indicating reactivity changes
      geom_point(
        #data = rc_changes,
        mapping = aes(
          y = (min_plot_value + margin_value) / 2,
          color = condition,
          alpha = reactivity_change
        ),
        size = 1.5,
        show.legend = FALSE,
        shape = 22,
        position = position_dodge(width = 0.8),
        stroke = POINT_STROKE *1.5
      ) +
      # lines separating residues
      geom_vline(
        xintercept = 1.5:length(unique(df$residue)),
        size = LINE_WIDTH,
        linetype = "solid",
        color = "grey90"
      ) + 
      # margin line
      geom_hline(
        yintercept = margin_value,
        size = LINE_WIDTH,
        linetype = "solid",
        color = "black"
        )  +
      labs(y = substitute(paste('log'[2],"(FC from ", m, ")"),list(m = control_condition)),
           x = "",
           fill = "Reactivity") +
      scale_fill_manual(values = cols) +
      scale_alpha_discrete(range = c(0,1)) +
      scale_color_manual(values = cols) +
      scale_shape_manual(values = c("Reactivity" = 21,"Protein expression" = 24)) +
      scale_y_continuous(limits = c(min_plot_value, max_plot_value), expand = expansion(mult = c(0, 0)))
    
    print(plot)
    
    save_plot(
      paste0(results_dir, "/05_boxplots/", paste(proteins_to_plot, collapse = "_")),
      width = wd,
      height = ht
    )
    
    return(df)
  }

'/Users/henrysanford/dev/test_data/macrophage/01_whole_proteome/03_results/3_reps/03_combined_files/1/07_combfiles_percctrl_wp_min_2_replicates.csv' %>%
  read_csv(show_col_types = FALSE) -> wp_table

'/Users/henrysanford/dev/test_data/macrophage/02_reactivity/03_rc_analysis/20250225_3reps/04_combined_files/2/07_combfiles_cysaggr_percctrl_rc_min_2_replicates.csv' %>%
  read_csv(show_col_types = FALSE) -> rc_table

 expression_and_reactivity_boxplot_all_conditions(
  whole_proteome_table = wp_table,
  reactivity_table = rc_table,
  proteins_to_plot = c("TBCK", "PPP1R21"),
  control_condition = "M0",
  ht = 2,
  wd = 6
) 
```


```{r, fig.width=2.5, fig.height=2}
expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("ARAP1"),
control_condition = "M0",
ht = 2,
wd = 3)

expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("ARHGEF1"),
control_condition = "M0",
ht = 2,
wd = 2.5)

expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("CDC42"),
control_condition = "M0",
ht = 2,
wd =2)

expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("AP3D1"),
control_condition = "M0",
ht = 2,
wd =2)

expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("PIP5K1A"),
control_condition = "M0",
ht = 2,
wd =2)

expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("RAB12"),
control_condition = "M0",
ht = 2,
wd =2)


full_df %>% filter(protein == "RAB12", condition == "TLR7")
```


```{r, fig.height=2, fig.width=2.2}
expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("GSR"),
control_condition = "M0",
ht = 2,
wd = 2.2)

```


```{r, fig.height=2, fig.width=10}

expression_and_reactivity_boxplot_all_conditions(
whole_proteome_table = wp_table,
reactivity_table = rc_table,
proteins_to_plot = c("PLEC"),
control_condition = "M0",
ht = 2,
wd = 7)

full_df %>%
  arrange(desc(rc_n_peptides)) %>%
  filter(reactivity_change)
```


```{r, fig.height=2.1, fig.width=3.2}
expression_boxplot_all_conditions <-
  function(condition_to_display,
           control_condition,
           proteins_to_plot,
           whole_proteome_table,
           ht,
           wd) {
    whole_proteome_table %>%
      # select columns needed for plot
      select(starts_with(c(
        conditions, c("protein", control_condition)
      ))) %>%
      filter(protein %in% proteins_to_plot) %>%
      # concat protein and residue
      column_to_rownames("protein") -> si_matrix
    print(si_matrix)
    
    # calculate fold change to median of control condition
    fold_change = si_matrix / apply(si_matrix[, grepl(control_condition, colnames(si_matrix))], 1, median, na.rm =
                                      T)
    
    
    fold_change %>%
      rownames_to_column("protein") %>%
      # pivot to long format for plotting
      pivot_longer(cols = colnames(.)[-1],
                   names_to = "condition_replicate",
                   values_to = "channel_ratio") %>%
      # drop missing values
      drop_na() %>%
      separate_wider_delim(
        cols = c("condition_replicate"),
        delim = "_",
        names = c("condition", "technical_replicate", "experiment"),
        too_many = "drop"
      ) %>%
      filter(protein %in% proteins_to_plot) -> wp_df
    wp_df$median_rc <- 0
    wp_df$residue <- "Expression"
    wp_df$regulation <- "Protein expression"
    wp_df$residue_loc <- 0
    
    wp_df %>%
      group_by(protein) %>%
      mutate(median_wp = median(channel_ratio)) -> wp_df
    wp_df %>% select(protein, median_wp) -> wp_stats
    
    wp_df$reactivity_change = FALSE
    wp_df$LFC_tmt_abpp = 0
    wp_df$wp_rc_ratio = 0
    wp_df$regulation = "Protein expression"
    
    wp_df -> df
    df$protein <- factor(df$protein, levels = proteins_to_plot)
    max_plot_value <- max(log2(df$channel_ratio))
    df %>% filter(condition != "TLR9") %>% mutate(condition = factor(condition, levels = c(control_condition, conditions))) ->  df
    
    breaks <- c(-1,0, 1)
    df %>%
      unique() %>%
      ggplot(aes(
        x = protein,
        fill = condition,
        y = log2(channel_ratio),
        shape = regulation
      )) +
      geom_hline(
        yintercept = breaks,
        size = LINE_WIDTH,
        linetype = "solid",
        color = "grey90"
      ) +
      geom_boxplot(
        outliers = FALSE,
        width = 0.5,
        lwd = LINE_WIDTH,
        show.legend = FALSE,
        alpha = 0.8,
        position = position_dodge(width = 0.8)
      ) +
      geom_beeswarm(
        #corral.width = 1.2,
        #cex = 2,
        col = "black",
        stroke = POINT_STROKE,
        show.legend = FALSE,
        size = 1,
        alpha = 0.9,
        dodge.width = 0.8
      ) +
      my_theme() +
      theme(
        plot.title = element_text(
          size = 6,
          color = "black",
          family = "Helvetica",
          hjust = 0.5
        ),
        strip.background = element_blank(),
        panel.spacing = unit(2, "mm"),
        #panel.border = element_blank(),
        strip.text = element_text(
          size = 8,
          family = "Helvetica",
          color = "black"
        ),
        #axis.line.x = AXIS_LINE,
        #axis.line.y = AXIS_LINE,
        strip.clip = "off"
      ) +
      labs(y = substitute(paste('log'[2],
                                "(FC from ", m, ")"),
                          list(m = control_condition)),
           x = "",
           fill = "Reactivity") +
      scale_fill_manual(values = cols) +
      scale_alpha_discrete(range = c(0,1)) +
      scale_color_manual(values = cols) +
      scale_shape_manual(values = c(
        "Higher" = 21,
        "Unchanged" = 21,
        "Lower" = 21,
        "Protein expression" = 24
        
      )) +
      geom_vline(
        xintercept = 1.5:length(unique(df$protein)),
        size = LINE_WIDTH,
        linetype = "solid",
        color = "grey90"
      ) +
      scale_y_continuous(breaks = breaks, limits = c(-1.5, 1.5))-> plot
    file_name <- paste(proteins_to_plot, collapse = "_")
    
    print(plot)
    
    save_plot(
      paste0(results_dir, "/05_boxplots/", file_name),
      width = wd,
      height = ht
    )
  }
ferry_proteins <- c(
  "TBCK",
  "PPP1R21",
  "FERRY3",
  "CRYZL1",
  "GATD1"
)
expression_boxplot_all_conditions(
  whole_proteome_table = wp_table,
  proteins_to_plot = ferry_proteins,
  control_condition = "M0",
  ht = 2.1 - 0.2255,
  wd = 3.2
)
```


## reactivity change barplot

```{r}
reactivity_change_barplot <- function(
    cys_df,
    highlight_uniprots,
    highlight_name,
    height = 2,
    width = 3,
    lower_rc_max = 80,
    higher_rc_max = 500,
    highlight_color = "#28ae80",
    rotate_x_labels = FALSE
){
x_axis_exand <- 0.2
colors <- c(
  "Other" = "grey90"
)
colors[highlight_name] = highlight_color
cys_df$direction_of_reactivity_change <- factor(cys_df$direction_of_reactivity_change,c("Higher", "Lower"))

if (length(highlight_uniprots) > 0){
cys_df %>%
  mutate(group = case_when(uniprot %in% highlight_uniprots ~ highlight_name,
                                   TRUE ~ "Other")) -> cys_df
}
cys_df$group <- factor(cys_df$group, levels = names(colors))
cys_df$condition <- factor(cys_df$condition, levels = conditions)
cys_df %>%
  ggplot(aes(x = condition, fill = group)) -> mito_plot
mito_plot +
    geom_bar(
           stat="count", show.legend = TRUE, width = 0.7, col = "black", size = LINE_WIDTH) + 
  geom_text(
    data = cys_df %>% filter(direction_of_reactivity_change == "Higher"),
    stat = "count", 
            inherit.aes = FALSE,
            mapping = aes(x = condition, label = ..count..), 
            vjust = -0.1, 
            size = FONT_SIZE, 
            size.unit = "pt") + 
    geom_text(
    data = cys_df %>% filter(direction_of_reactivity_change == "Lower"),
    stat = "count", 
            inherit.aes = FALSE,
            mapping = aes(x = condition, label = ..count..), 
            vjust = 1.1, 
            size = FONT_SIZE, 
            size.unit = "pt") + 
  facet_grid("direction_of_reactivity_change", scales = "free_y", space = "free") + 
  guides(fill = guide_legend(ncol =1)) +  
  legend_theme() +
  my_theme() +
  theme(
    strip.background = element_blank(),
    panel.spacing = unit(0, "mm"),
  ) +
  xlab("") +
  scale_fill_viridis_d(direction = -1) +
  scale_fill_discrete(type = colors) +
  ylab("number of reactivity changes") + 
  scale_y_continuous(expand = c(0,0)) +
  facetted_pos_scales(
    y = list(
      scale_y_continuous(expand = expansion(mult = c(0,x_axis_exand)), limits = c(0, higher_rc_max)),
      scale_y_continuous(expand = expansion(mult = c(x_axis_exand,0)), trans = "reverse", limits = c(lower_rc_max,0))
      )
    ) -> plt
  if (rotate_x_labels) {
    plt <- plt +
      theme(
        axis.text.x = element_text(
          color = "black",
          angle = 45,
          size = FONT_SIZE,
          family = "Arial",
          vjust = 1,
          hjust = 1
        )
      )
  }
print(plt)
save_plot(
  fn = paste0(results_dir, "/07_barplots/", highlight_name, "_rc_barplot"),
  width = width,
  height = height
)
}


```

```{r, fig.width=3.1, fig.height=1.7}

protein_list_dir <-
  '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/00_reference lists/GO_term_lists/'

for (list_dir in list.dirs(protein_list_dir, recursive = FALSE)) {
  list_name <- str_split(list_dir, "//")[[1]][2]
  if (list_name != "Vesicle-mediated transport"){
    next
  }
  # MSigDB 2026 lists (GO_updates.md migration) are gene-symbol-only; map to the
  # UniProt accessions measured in the data so the uniprot-based highlighting is unchanged.
  list_genes <- paste0(list_dir, "/02_genes_matching_query.csv") %>%
    read_csv(show_col_types = FALSE) %>%
    pull(genes)
  list_uniprots <- full_df %>%
    filter(protein %in% list_genes) %>%
    pull(uniprot) %>%
    unique()

  reactivity_change_barplot(
    cys_df = full_df %>% filter(reactivity_change),
    highlight_uniprots = list_uniprots,
    highlight_name = list_name,
    lower_rc_max = 170,
    higher_rc_max = 390,
    height =1.7, 
    width = 3.1,
    highlight_color = "#237D41",
    rotate_x_labels = TRUE
  )
}
```
```{r, fig.width=2.8, fig.height=2}
hn <- "Phosphorylation\nsite on peptide"
"/Users/henrysanford/dev/test_data/macrophage/02_reactivity/03_rc_analysis/20250225_3reps/05_reactivity_changes/output_median_for_5+_peptides/cysteine_reactivity_changes.csv" %>%
  read_csv() %>% 
  filter(condition != "TLR9") %>%
  mutate(group = case_when(phosphorylation == TRUE ~ hn, TRUE ~ "Other")) -> df


reactivity_change_barplot(
  cys_df = df,
  highlight_uniprots = c(),
  highlight_name = hn,
    lower_rc_max = 130,
    higher_rc_max = 400,
    height =2, 
    width = 2.8,
    highlight_color = "#5CBED7",
    rotate_x_labels = TRUE
)
```




```{r}

paste0("/Users/henrysanford/dev/test_data/macrophage/02_reactivity/03_rc_analysis/20250225_3reps/05_reactivity_changes/output_median_for_5+_peptides/cysteine_reactivity_changes.csv") %>%
  read_csv(show_col_types = FALSE) %>% filter(condition != "TLR9") -> cys_df

  reactivity_change_barplot(
    cys_df = cys_df,
    highlight_uniprots = list_uniprots,
    highlight_name = list_name,
    lower_rc_max = 130,
    higher_rc_max = 420,
    height =2, 
    width = 3.1,
    highlight_color = "#237D41",
    rotate_x_labels = TRUE
  )
```



```{r, fig.height=4, fig.width=22}

df %>% 
  filter(phosphorylation == TRUE, condition != "TLR9") %>%
  mutate(reactivity_change_int = 1,
         name = paste0(protein, "_", residue)) %>%
  select(name, condition, reactivity_change_int) %>% 
  pivot_wider(id_cols = c("name"), names_from = condition, values_from = reactivity_change_int) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  column_to_rownames("name") %>%
  t() -> mat

cn = colnames(mat)
pdf("phosphosites.pdf", width = 22, height = 4.5)
Heatmap(mat, show_column_names = FALSE, 
       row_dend_width = unit(17, "mm"),
        bottom_annotation = HeatmapAnnotation(
        text = anno_text(cn, rot = 45, offset = unit(1, "npc"), just = "right"),
        annotation_height = max_text_width(cn)
    ), col = colorRamp2(c(0,1), c("white", "#5CBED7")),
    rect_gp = gpar(col = "grey", lwd = LINE_WIDTH * 2),show_heatmap_legend = FALSE) -> ht

print(ht)
dev.off()
print(ht)
```


### Phosphoproteomics volcano plot
```{r}
volcano_data
```


```{r, fig.width=2, fig.height=2}
library(ggbreak)
results_dir <- "/Users/henrysanford/dev/test_data/phospho/tiff_3reps/output/3reps/05_results"
paste0(
  results_dir, "/volcano_plot/volcano_data.csv"
) %>% 
  read_csv(show_col_types = FALSE) %>%
  mutate(residue_label = ID) -> volcano_data
sig_up_col <- "indianred2"
sig_down_col <- "steelblue2"
unchanged_col <- "lightgrey"

alphas_two = c("Significant Down" = 0.6, 
  "Significant Up" = 0.6, 
  "Not Significant Down" = 0.2, 
  "Not Significant Up" = 0.2, 
  "Not Significant" = 0.2,
  "Significant but <1.5 FC" = 0.2)

cols_three = c("Significant Down" = sig_down_col, 
  "Significant Up" = sig_up_col, 
  "Not Significant Down" = sig_down_col, 
  "Not Significant Up" = sig_up_col, 
  "Not Significant" = unchanged_col,
  "Significant but <1.5 FC" = unchanged_col)

volcano_data$residue_label <- gsub(";", ",", volcano_data$residue_label)

volcano_data$`Regulation_phospho - TLR4 vs. M0 (5634 Residues)` <- factor(
  volcano_data$`Regulation_phospho - TLR4 vs. M0 (5634 Residues)`, 
  levels = rev(names(cols_three))
  )

volcano_data %>% 
  filter(`ID` %in% c("STAT4_Y693", "NFKB2_S872", "NFKB1_S937,S941","NFKB1_S907")) -> label_data
fc_label = expression('phosphorylation, log'[2]*'(TLR4/M0)')
p_value_label = expression('-log'[10]*'(p-value)')
volcano_data %>%
  arrange(`Regulation_phospho - TLR4 vs. M0 (5634 Residues)`) %>%
  ggplot(
    aes(x = `log2_FC_phospho - TLR4 vs. M0 (5634 Residues)`,
        y = `-log10_pval_phospho - TLR4 vs. M0 (5634 Residues)`,
        label = residue_label,
        fill = `Regulation_phospho - TLR4 vs. M0 (5634 Residues)`,
        alpha = `Regulation_phospho - TLR4 vs. M0 (5634 Residues)`)
    ) + 
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size=LINE_WIDTH) +
  geom_vline(xintercept = c(-FC_CUTOFF, FC_CUTOFF), linetype = "dashed", size=LINE_WIDTH) + 
  geom_point(
    shape = 21,
    size = 1,
    stroke = POINT_STROKE /2,
    show.legend = FALSE
  ) + 
  geom_text_repel(data = label_data,
                  mapping = aes(color = `Regulation_phospho - TLR4 vs. M0 (5634 Residues)`),
                  size = 1.940,
                  nudge_x = 8,
                  box.padding = 1,
                  color = "black",
                  min.segment.length = 0,
                  segment.size = LINE_WIDTH,
                  show.legend = FALSE,  alpha = 1, max.overlaps = Inf) +
  my_theme() + 
  scale_x_continuous(limits = c(-4,18)) +
  #scale_x_break(breaks = c(7, 16), space = 0.01, scales = 0.1) +
  scale_fill_manual(values = cols_three) + 
  scale_color_manual(values = cols_three) + 
  scale_alpha_manual(values = alphas_two) +
  xlab(fc_label) +
  ylab(p_value_label)

ggsave(paste0(results_dir, "/volcano_plot/volcano_plot.svg"), width = 2, height = 2)
ggsave(paste0(results_dir, "/volcano_plot/volcano_plot.png"), width = 2, height = 2)
```

```{r, fig.width=2, fig.height=2}
volcano_data %>%
  arrange(`Regulation_phospho - TLR4 vs. M0 (5634 Residues)`) %>%
  mutate(x_range = ifelse(`log2_FC_phospho - TLR4 vs. M0 (5634 Residues)` <= 8, "≤ 8", "> 8")) %>%
  mutate(x_range = fct_relevel(x_range, "> 8", "≤ 8")) %>%  # Explicit ordering
  arrange(x_range) %>%
  ggplot(
    aes(x = `log2_FC_phospho - TLR4 vs. M0 (5634 Residues)`,
        y = `-log10_pval_phospho - TLR4 vs. M0 (5634 Residues)`,
        label = residue_label,
        fill = `Regulation_phospho - TLR4 vs. M0 (5634 Residues)`,
        alpha = `Regulation_phospho - TLR4 vs. M0 (5634 Residues)`)
    ) + 
  geom_point(
    shape = 21,
    size = 1,
    stroke = POINT_STROKE /2,
    show.legend = FALSE
  ) + 
  geom_text_repel(data = label_data %>% 
                    mutate(x_range = ifelse(`log2_FC_phospho - TLR4 vs. M0 (5634 Residues)` <= 8, "≤ 8", "> 8")) %>%
                    mutate(x_range = fct_relevel(x_range, "> 8", "≤ 8")),
                  mapping = aes(color = `Regulation_phospho - TLR4 vs. M0 (5634 Residues)`),
                  size = 1.940,
                  nudge_x = 8,
                  box.padding = 1,
                  color = "black",
                  min.segment.length = 0,
                  segment.size = LINE_WIDTH,
                  show.legend = FALSE,  alpha = 1, max.overlaps = Inf) +
  facet_grid(cols = vars(x_range), scales = "free_x", space = "free") +
  geom_hline(data = data.frame(x_range = "≤ 8", y = -log10(0.05)), aes(yintercept = y), linetype = "dashed", size = LINE_WIDTH, inherit.aes = FALSE) +
  geom_vline(data = data.frame(x_range = "≤ 8", x = c(-FC_CUTOFF, FC_CUTOFF)), aes(xintercept = x), linetype = "dashed", size = LINE_WIDTH, inherit.aes = FALSE) +
  my_theme() + 
  theme(strip.text = element_blank(),
        panel.spacing = unit(0, "lines")) +
  scale_fill_manual(values = cols_three) + 
  scale_color_manual(values = cols_three) + 
  scale_alpha_manual(values = alphas_two) +
  xlab(fc_label) +
  ylab(p_value_label)
```

### Heatmap

```{r, fig.height=1.5, fig.width=5.4}

GEF <- c("GBF1","ARHGEF7","DNMBP","DOCK4","DOCK7","TIAM1","RAB3IL1","DENND4A","DENND4C","GAPVD1","RAPGEF1","AKAP13","ARHGEF1","TRIO","DOCK8","PREX1")
GAP <- c("AGFG2","GIT1","ARAP1","TBC1D15","TBC1D2","TBC1D8","TBC1D9B","FLCN","SIPA1","RASA3","ARHGAP31","ARHGAP4","ARHGAP21","MYO9B","ARHGAP30")
GTPase <- c("CDC42", "EFL1","SEPTIN9","RAB12")

full_df %>%
  mutate(gef_gap = case_when(protein %in% GTPase ~ "GTPase", protein %in% GEF ~ "GEF", TRUE ~ "GAP"), condition = factor(condition, levels = rev(conditions))) %>%
  arrange(condition) %>%
  filter(
    protein %in% c(GEF, GAP,GTPase)
  ) %>% 
  filter(direction_of_reactivity_change != "unchanged") %>%
  group_by(protein, condition, gef_gap) %>%
  summarise(
    direction_of_reactivity_change = paste(unique(direction_of_reactivity_change),  collapse = ", ")
  ) %>%
  mutate(
    direction_of_reactivity_change = case_when(
      (grepl(",",direction_of_reactivity_change)) ~ "Both",
      TRUE ~ direction_of_reactivity_change
      )
  ) %>%
  ggplot(aes(x = protein, y = condition, fill = direction_of_reactivity_change)) +
  geom_tile(linewidth = 0.25, color = "white", show.legend = FALSE) +
      scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  my_theme() + 
        theme(
        axis.text.x = element_text(
          color = "black",
          angle = 45,
          size = 6,
          family = "Arial",
          vjust = 1,
          hjust = 1
        ),
                panel.spacing = unit(1, "lines"),
        strip.background = element_blank(),
        strip.clip = "off"
      ) +
  scale_fill_manual(values = c("Higher" = SIG_UP_COL, "Lower" = SIG_DOWN_COL, "Both" = "purple"))+ 
  facet_grid(cols = vars(gef_gap), scales = "free_x", space = "free")
save_plot(
  paste0(results_dir, "/08_heatmap/gef_gap.svg"), width = 4.8, height = 1.5
)
```




# Alluvial plot

```{r, fig.height=1.9, fig.width=5}

cols <- c(
  "TLR1-2"= "#EC2427",
   "M0" = "#AAAAAA",
  "TLR3" = "#F58420",
  "TLR4" = "#FFCC31",
  "TLR7" = "#4869B2",
  "TLR8" = "#5CBED7",
  "TLR9" = "#237D41",
  "STING" = "#742D16"
)

# Create alluvial plot
create_alluvial <- function(df, columns) {
  
  # Count flows
  flow_data <- df %>%
    group_by(across(all_of(columns))) %>%
    summarise(n = n(), .groups = "drop")
  
  
  # Create the plot
  p <- ggplot(flow_data,
              aes(axis1 = .data[[columns[1]]],
                  axis2 = .data[[columns[3]]],
                  axis3 = .data[[columns[4]]],
                  axis4 = .data[[columns[2]]],
                  axis5 = .data[[columns[5]]],
                  y = n)) +
    geom_alluvium(
                  aes(fill = .data[[columns[1]]]), 
                  alpha = 0.7,
                  width = 2/3) +
    geom_stratum(
      aes(fill = after_stat(stratum)),
      width = 2/3, 
                 color = "black",
                 linewidth = 0.25 / 2) +
    geom_text(stat = "stratum", 
              aes(label = paste(after_stat(stratum), after_stat(count))),
              size = 5.5 / .pt,  # Convert pt to mm for ggplot
              family = "Arial") +
    theme_minimal() +
    theme(
      text = element_text(family = "Arial", size = 6),
      axis.text.y = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      legend.position = "none"
    ) +scale_fill_manual(values = rep(lighten(unname(cols), amount = 0.25), 12))
  
  return(p)
}

flags <- c("reactive" = TRUE, "unreactive" = FALSE)
# Load  data
df <- read_csv("/Users/henrysanford/Downloads/formatted_cross_ref.csv")
  
columns_to_plot <- c(
  "Complex",
  "Depth of Conservation (# of Organisms)",
  "pPSE",
  "structure_group",
  "am_class"
)
for (name in names(flags)){

  # Create plot
  p <- create_alluvial(df |> filter(reactivity_change == flags[name]), columns_to_plot)
  
  # Save as PDF (3 inches tall x 6 inches wide)
  ggsave(
    paste0(name, "_alluvial_plot.pdf"),
    plot = p,
    width = 5,
    height = 1.9,
    units = "in",
    device = cairo_pdf  # Better font handling
  )
  
}




```

```{r, fig.height=2, fig.width=1.5}
df <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/01_data_analysis_folders/06_macrophage/macrophage/reactivity/species.csv' %>%
  read_csv(show_col_types = FALSE) %>%
  select(uniprot, residue, mm10, `D. melanogaster`, `C. elegans`, `S. cerevisiae`) %>%
  unique()

df %>%
  pivot_longer(cols = c(mm10, `D. melanogaster`, `C. elegans`, `S. cerevisiae`)) %>%
  drop_na() %>%
  mutate(value = factor(value, levels = c(0,1)), 
         name = case_when(name == "mm10" ~ "M. musculus", TRUE ~ name),
         name = factor(name, levels = c("M. musculus","D. melanogaster", "C. elegans", "S. cerevisiae" ))) %>%
  arrange(name) %>%
  group_by(name, value) %>%
  summarise(count = n(), .groups = "drop_last") %>%
  mutate(percentage = count / sum(count) * 100) %>%
  filter(value == 1) %>%
  ggplot(aes(x = name, y = percentage)) +
  geom_col(width = 0.7, fill = cols[["TLR8"]], color = "black", size = LINE_WIDTH) +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), size = FONT_SIZE_MM ,vjust = -0.5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 100)) +
  my_theme() +
       theme(axis.text.x = element_text(
        color="black",angle=45, size=FONT_SIZE, family="Arial", vjust = 1, hjust=1)) + labs(x = NULL, y = "percent cysteine conservation")

save_plot("conservation", height = 2, width = 1.5)
```

# GO term correlaiton

```{r, fig.height=7.9, fig.width=4.5}
col_location <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Vinogradova Laboratory/Henry_data processing/01_data_analysis_folders/06_macrophage/macrophage/whole_proteome/'
df <- NULL
for (cond in conditions){
  for (color in c("violet", "teal-yellow")){
    for (dir in c("Up", "Down")){ # not including "both" category
      fn <- paste0(col_location, cond,"_", color,"_", dir, ".csv")
      if(file.exists(fn)){
        df <- fn %>%
          read_csv(show_col_types = FALSE) %>%
          mutate(
            direction = dir,
            condition = cond,
            col = color,
            term = Term  # standard schema from go_enrich() (macrophage_figure_setting.R)
          ) %>%
          filter(adjusted_p_value < 0.05, n_proteins > 5, ontology == "BP")  %>%
          rbind(df)
        print(nrow(df))
      }
    }
  }
}

df %>%
  mutate(
    condition = factor(condition, levels = conditions),
    dir = factor(direction, levels = c( "Down", "Up"))
  ) %>%
  arrange(direction) %>%
  group_by(col, condition, direction, term) %>%
  arrange(direction) %>%
  mutate(
    term = fct_inorder(str_wrap(term, 50)),
    col = case_when(
      col == "teal-yellow" ~ "teal-yellow (antidirectional)",
      TRUE ~ "violet (RNA only)"
    )
  ) %>%
  ggplot(aes(
    x = direction,
    y = term,
    size = n_proteins,
    fill = -log10(adjusted_p_value)
  )) +
  geom_point(shape = 21, stroke = LINE_WIDTH) +
  my_theme()  +
  labs(
    x = NULL,
    y = NULL,
    fill = expression("-log"[10] * "(FDR)"),
    size = "proteins"
  ) +
  facet_grid(
    rows = vars(col),
    cols = vars(condition),
    space = "free",
    scales = "free"
  ) +
  theme(
    strip.background = element_blank(),
    strip.clip = "off",
    panel.grid.major = element_line(color = "grey90", linewidth = LINE_WIDTH)
  ) +
  scale_fill_gradient(low= "#b3cde3", high = "#810f7c") +
  theme(
    axis.text.x = element_text(
      color = "black",
      angle = 45,
      size = FONT_SIZE,
      family = FONT_FAMILY,
      vjust = 1,
      hjust = 1
    ),
    axis.text.y = element_text(
      color = "black",
      family = FONT_FAMILY,
      size = 5
    ),
    strip.text = element_text(size = FONT_SIZE, color = "black", family = FONT_FAMILY),
    panel.spacing = unit(0, "mm")
  )+ scale_y_discrete(limits = rev) + 
  scale_size_continuous(range = c(1, 4))


ggsave('/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/Figure 2_Unenriched proteomics analysis/Panels/wp_vs_rnaseq_go_term/rna_vs_wp_correlation_go_term_analysis_BP.pdf', height = 7.9, width = 4.5)
```

# GO:BP enrichment of TLR4 reactivity changes

```{r, fig.height=2, fig.width=3.19}
go_out_dir <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 2_Reactivity_part1_GTPase regulation/Panels/go_enrichment"

# TLR4 proteins (gene symbols) with a significant reactivity change, from the
# reactivity_polars.ipynb output.
tlr4_reactive <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/02_reactivity/03_rc_analysis/reactivity_output/20251023_ratio_directionality/median_for_5+_peptides/output/reactivity_changes_long_format.csv' |>
  read_csv(show_col_types = FALSE) |>
  filter(reactivity_change, condition == "TLR4") |>
  pull(protein) |>
  unique()

# Centralized Fisher-exact GO:BP over-representation (go_enrich() in
# macrophage_figure_setting.R: clusterProfiler::enricher on MSigDB 2026 GO, simplified).
# Background = NULL (all annotated genes), matching the original enrichGO default universe.
go_bp_tlr4 <- go_enrich(tlr4_reactive, ont = "BP", background = NULL, simplify_cutoff = 0.5)
write_csv(go_bp_tlr4, file.path(go_out_dir, "go_results.csv"))

go_bp_tlr4 |>
  filter(adjusted_p_value < 0.05, n_proteins > 5) |>
  slice_min(adjusted_p_value, n = 20) |>
  go_dotplot(size_range = c(1, 3), title = "GO:BP enrichment analysis (TLR4)")

ggsave(file.path(go_out_dir, "go_enrichment.pdf"), width = 230/72, height = 144/72, device = cairo_pdf)
```

# Cross-omics GO enrichment of the venn terms

Complements the GO-term venn diagrams in reactivity_polars.ipynb (which show raw overlap of a
term's genes with each omic's changes): here we test whether each term is statistically enriched
among each omic's full set of significant changes, via go_enrich() (Fisher-exact over MSigDB C5 GO).
Term gene sets therefore come from msigdbr, not the venn GMT/query lists, so counts differ slightly.

```{r}
prefix <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/01_Data Analysis/01_Proteomics/"
dirs <- c("Significant Up", "Significant Down")

# Changed-protein (gene symbol) set per omic, all TLR4 vs. M0. Mirrors omic_change_sets() in
# reactivity_polars.ipynb; reactivity is restricted to condition == "TLR4" to match the
# standalone GO:BP enrichment above (go_bp_tlr4).
omic_changes <- list(
  "Reactivity (TLR4)" = read_csv(paste0(prefix, "02_reactivity/03_rc_analysis/reactivity_output/20251023_ratio_directionality/median_for_5+_peptides/output/reactivity_changes_long_format.csv"),
                          col_select = c(protein, condition, reactivity_change), show_col_types = FALSE) |>
    filter(reactivity_change, condition == "TLR4") |> pull(protein) |> unique(),
  # Upregulated-only, matching the panel F "phosphosites up in TLR4" set in
  # phosphoproteomics/visualization.Rmd (log2FC > 1 & p < 0.05, unique proteins).
  "Phospho (TLR4)" = read_csv(paste0(prefix, "05_phosphoproteomics/processed_results/expression-normalized_phosphorylation_table.csv"),
                              col_select = c(protein, `log2_FC_phospho - TLR4 vs. M0 (9591 Residues)`, `-log10_pval_phospho - TLR4 vs. M0 (9591 Residues)`), show_col_types = FALSE) |>
    filter(`log2_FC_phospho - TLR4 vs. M0 (9591 Residues)` > 1,
           `-log10_pval_phospho - TLR4 vs. M0 (9591 Residues)` > -log10(0.05)) |> pull(protein) |> unique(),
  "Expression (TLR4)" = read_csv(paste0(prefix, "01_unenriched proteomics/03_results/20250219_3reps/04_results/volcano_plots/volcano_data_long_format.csv"),
                                 col_select = c(protein, condition, Regulation), show_col_types = FALSE) |>
    filter(condition == "TLR4", Regulation %in% dirs) |> pull(protein) |> unique()
)

# Terms of interest. max_gs = the maxGSSize used for that term's enrichment: endosomal
# transport / GTPase use 500 (identical universe to go_bp_tlr4, so adjusted_p_value matches
# that analysis exactly; endocytosis was dropped for exceeding it). Vesicle-mediated transport
# has 1643 genes, so it needs a larger cap (2000) and is enriched in its own run below -- its
# FDR is therefore BH-adjusted over a larger term universe than the 500-cap terms.
terms <- tribble(
  ~Term,                        ~GO_id,       ~ont,  ~max_gs,
  "Endosomal transport",        "GO:0016197", "BP",   500L,
  "GTPase activity",            "GO:0003924", "MF",   500L,
  "Vesicle-mediated transport", "GO:0016192", "BP",  2000L
)

# One go_enrich() per (omic, ontology, maxGSSize) group, keeping only that group's target
# term(s). Background = all annotated genes; cutoffs relaxed + no simplify so non-significant
# terms are kept (written to CSV so any term can be plotted later).
enr <- purrr::imap_dfr(omic_changes, function(genes, omic) {
  purrr::pmap_dfr(distinct(terms, ont, max_gs), function(ont, max_gs) {
    keep <- terms$GO_id[terms$ont == ont & terms$max_gs == max_gs]
    go_enrich(genes, ont = ont, background = NULL, simplify_cutoff = NULL,
              pvalueCutoff = 1, qvalueCutoff = 1, minGSSize = 10, maxGSSize = max_gs) |>
      filter(GO_id %in% keep)
  }) |> mutate(omic = omic)
})

# Complete the omic x term grid; terms with zero overlap get p = 1, n = 0.
res <- tidyr::expand_grid(GO_id = terms$GO_id, omic = names(omic_changes)) |>
  left_join(terms, by = "GO_id") |>
  left_join(select(enr, GO_id, omic, p_value, adjusted_p_value, n_proteins, Genes),
            by = c("GO_id", "omic")) |>
  mutate(p_value = coalesce(p_value, 1), adjusted_p_value = coalesce(adjusted_p_value, 1),
         n_proteins = coalesce(n_proteins, 0L),
         omic = factor(omic, levels = names(omic_changes)))

# Save alongside the GO-term venn diagrams (reactivity_polars.ipynb) these terms belong to.
# Full results (all tested terms, non-significant included) + the targeted subset, so any
# term can be re-plotted from the CSV. Plotting itself is the next chunk.
venn_go_dir <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 3_Reactivity_part2_endocytosis/Panels/go_term_venn_diagram"
write_csv(enr, file.path(venn_go_dir, "cross_omics_go_enrichment_all_terms.csv"))
write_csv(res, file.path(venn_go_dir, "cross_omics_go_enrichment.csv"))
```
DISC1
```{r, fig.height=2.4, fig.width=7.5}
# Disease/phenotype associations for the FERRY-complex reactive-cysteine genes
# (PPP1R21, TBCK) plus DISC1, pulled from the Open Targets Platform by
# src/opentargets_utils.py (re-run that script to refresh the CSV). x = overall
# association score, bubble size = number of supporting evidence datatypes, one
# facet per gene -- same dot-plot style as the GO panels (go_dotplot).
pheno_df <- read_csv("../reference_dbs/phenotype_associations.csv", show_col_types = FALSE) |>
  # Wrap long disease names so the y-axis labels stay legible.
  mutate(disease = stringr::str_wrap(disease, width = 30))

go_dotplot(
  pheno_df,
  term_col     = "disease",
  fdr_col      = "score",
  size_col     = "n_evidence",
  facet        = "gene",
  facet_scales = "free_y",  # shared x so scores are comparable across genes
  x_transform  = identity,
  size_range   = c(1, 3),
  size_name    = "evidence",
  xlab         = "Open Targets association score"
)

ggsave("/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 6/Panels/phenotype_association/phenotype_associations_dotplot.pdf",
       height = 2.4, width = 7.5)
```

```{r, fig.height=2.4, fig.width=7.5}
# Same panel, but restricted to nervous-system / psychiatric therapeutic areas before
# taking the top diseases per gene (phenotype_associations_neuro.csv, written by the same
# script with therapeutic_area_ids = NEURO_PSYCH_TA). This keeps DISC1's schizophrenia link
# on the panel instead of the pleiotropic GWAS hits that top its unfiltered raw-score list.
pheno_neuro_df <- read_csv("../reference_dbs/phenotype_associations_neuro.csv", show_col_types = FALSE) |>
  mutate(disease = stringr::str_wrap(disease, width = 30))

go_dotplot(
  pheno_neuro_df,
  term_col     = "disease",
  fdr_col      = "score",
  size_col     = "n_evidence",
  facet        = "gene",
  facet_scales = "free_y",
  x_transform  = identity,
  size_range   = c(1, 3),
  size_name    = "evidence",
  xlab         = "Open Targets association score"
)

ggsave("/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 6/Panels/phenotype_association/phenotype_associations_neuro_dotplot.pdf",
       height = 2.4, width = 7.5)
```

```{r, fig.height=1.5, fig.width=2}
res |>
  filter(Term == "Endosomal transport") |>
  ggplot(aes(x = -log10(adjusted_p_value), y = omic, size = n_proteins)) + 
  facet_wrap(~Term) + 
  geom_point(shape =21, fill = "#B3CDE3", stroke = LINE_WIDTH) + 
  geom_vline(xintercept = -log10(0.01), linetype = "dashed", size = LINE_WIDTH) + 
  my_theme() +
  theme(strip.background = element_blank(), strip.clip = "off") + 
  labs(x = expression("-log"[10] * "(FDR)"), y = NULL, size = "proteins") + 
  scale_size_continuous(range = c(1, 3), breaks = c(20, 25,30, 35)) + 
  scale_x_continuous(expand = expansion(mult = c(0.2,0.2)))

ggsave(
  "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 3_Reactivity_part2_endocytosis/Panels/go_term_venn_diagram/Endosmal transport_cross_omics_enrichment.pdf", height = 1.5, width = 2
)


res |>
  arrange(p_value)
```

```{r, fig.height=1.5, fig.width=2}
# Same panel for GTPase activity (GO:0003924, MF). Shares the maxGSSize = 500 term
# universe with endosomal transport, so its FDR is comparable to that panel.
res |>
  filter(Term == "GTPase activity") |>
  ggplot(aes(x = -log10(adjusted_p_value), y = omic, size = n_proteins)) +
  facet_wrap(~Term) +
  geom_point(shape =21, fill = "#B3CDE3", stroke = LINE_WIDTH) +
  geom_vline(xintercept = -log10(0.01), linetype = "dashed", size = LINE_WIDTH) +
  my_theme() +
  theme(strip.background = element_blank(), strip.clip = "off") +
  labs(x = expression("-log"[10] * "(FDR)"), y = NULL, size = "proteins") +
  scale_size_continuous(range = c(1, 3), breaks = c(20, 25,30, 35)) +
  scale_x_continuous(expand = expansion(mult = c(0.2,0.2)))

ggsave(
  "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 3_Reactivity_part2_endocytosis/Panels/go_term_venn_diagram/GTPase activity_cross_omics_enrichment.pdf", height = 1.5, width = 2
)
```

```{r, fig.height=1.5, fig.width=2}
# Same panel for vesicle-mediated transport (GO:0016192). NB its FDR is BH-adjusted over the
# maxGSSize = 2000 term universe (the set has 1643 genes), unlike the 500-cap terms above.
res |>
  filter(Term == "Vesicle-mediated transport") |>
  ggplot(aes(x = -log10(adjusted_p_value), y = omic, size = n_proteins)) +
  facet_wrap(~Term) +
  geom_point(shape =21, fill = "#B3CDE3", stroke = LINE_WIDTH) +
  geom_vline(xintercept = -log10(0.01), linetype = "dashed", size = LINE_WIDTH) +
  my_theme() +
  theme(strip.background = element_blank(), strip.clip = "off") +
  labs(x = expression("-log"[10] * "(FDR)"), y = NULL, size = "proteins") +
  scale_size_continuous(range = c(1, 3)) +
  scale_x_continuous(expand = expansion(mult = c(0.2,0.2)))

ggsave(
  "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/6-figure format_EV/Figure 3_Reactivity_part2_endocytosis/Panels/go_term_venn_diagram/Vesicle-mediated transport_cross_omics_enrichment.pdf", height = 1.5, width = 2
)
```












# venn diagram enrichment

```{r}
dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Macrophage project/Manuscript 1/02_Figures/Figure 3_reactivity profiling broad/Panels/venn_diagram/'

venn_data <- paste0(dir, "venn_data.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(synonyms = protein)

hpa_data <- NULL
tissues_to_download <- c("Macrophages" = "macrophages", "T-cells" = "T-cells")
for (tissue in names(tissues_to_download)) {
  for (category in c("Cell+type+enhanced", "Cell+type+enriched", "Group+enriched")) {
    url <- paste0(
      "https://www.proteinatlas.org/search/cell_type_category_rna%3A",
      tissue,
      "%3B",
      category,
      "+AND+sort_by%3Atissue+specific+score?format=tsv"
    )
    hpa_data <- GET(url) %>%
      content(as = "text") %>%
      read_delim(show_col_types = FALSE) %>%
      mutate(
        tissue = tissues_to_download[tissue],
        hpa_category = gsub("\\+", " ", category)
      ) %>%
      rbind(hpa_data)

    Sys.sleep(0.1)
  }
}

df <- hpa_data %>%
  mutate(synonyms = paste0(Gene, ", ", `Gene synonym`)) %>%
  separate_rows(synonyms, sep = ", ") %>%
  filter(synonyms != "NA") %>%
  mutate(key = paste(synonyms, tissue)) %>%
  dplyr::select(synonyms, tissue, key, hpa_category, `RNA single cell type specificity`, `RNA single cell type specific nCPM`) %>%
  filter(synonyms %in% unique(venn_data$protein)) %>%
    group_by(synonyms) %>%
  summarise(
    tissue = case_when(
      all(c("T-cells", "macrophages") %in% tissue) ~ "enriched in both",
      "T-cells" %in% tissue ~ "enriched in T-cells",
      "macrophages" %in% tissue ~ "enriched macrophages"
    ), `RNA single cell type specificity`, `RNA single cell type specific nCPM`) %>%
  right_join(y = venn_data, by = "synonyms") %>%
  mutate(tissue = case_when(is.na(tissue) ~ "enriched in neither", TRUE ~ tissue))
unique(df$tissue)


df %>% write_csv(paste0(dir, "venn_data_annotated.csv"))
```

```{r, fig.width=3, fig.height=2}
library(ComplexUpset)
library(ggbreak)
TEXT_ELEMENT = element_text(size = FONT_SIZE,
                            color = "black",
                            family = FONT_FAMILY)

upset_plot_theme <-
  theme_classic() +
  theme(
    axis.text.y = TEXT_ELEMENT,
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_blank(),
    axis.title.y = TEXT_ELEMENT,
    panel.grid = element_blank(),
    axis.line.x = AXIS_LINE,
    axis.line.y = AXIS_LINE
  )
df %>%
  filter((`Macrophage` == 1) |
           (`Cell 2020` == 1)) %>%
  mutate(tissue = factor(tissue, levels = rev(c("enriched in both", "enriched in T-cells","enriched macrophages","enriched in neither")))) %>%
  upset(
    intersect = c("Macrophage", "Cell 2020"),
    matrix = (
      intersection_matrix(
        geom = geom_point(size = 1),
        segment = geom_segment(linewidth = LINE_WIDTH)
      )
    ),
    base_annotations =
      list(
        'Intersection size' =
          intersection_size(
            mapping = aes(fill = tissue),
            text = list(
              size = FONT_SIZE_MM,
              color = "black",
              family = FONT_FAMILY,
              vjust = -0.1,
              check_overlap = TRUE
            ),
            color = "black",
            size = LINE_WIDTH,
            width = 0.7
          ) + scale_fill_manual(values = rev(c(cols[["TLR1-2"]], cols[["TLR3"]], cols[["TLR7"]], cols[["M0"]]))) + theme(legend.text = TEXT_ELEMENT, legend.title = element_blank(), legend.key.size = unit(3, "mm")) + scale_y_break(c(30, 100), scales = 0.5)
      ),
    set_sizes = FALSE,
    stripes = "white",
    themes = list(
      intersections_matrix = upset_plot_theme + theme(
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.title.y = element_blank()
      ),
      `Intersection size` = upset_plot_theme + theme(
        axis.text.x = element_blank(),
        axis.title.y = TEXT_ELEMENT,
        axis.text.y = TEXT_ELEMENT,
        axis.line.y = AXIS_LINE,
        axis.ticks.x = element_blank()
      )
    )
  ) 

ggsave(paste0(dir, "enriched in cells upset_plot.svg"), width = 3, height =2 )
```






# Comparison with T cell data

```{r}
expression_and_reactivity_boxplot_all_conditions <-
  function(condition_to_display,
           control_condition,
           proteins_to_plot,
           whole_proteome_table,
           reactivity_table,
           ht,
           wd) {
    cysteine_seperator <- ", " # for aggregated cysteines
    
    si_matrix <- whole_proteome_table %>%
      # select columns needed for plot
      select(starts_with(c(
        conditions, c("protein", control_condition)
      ))) %>%
      filter(protein %in% proteins_to_plot) %>%
      column_to_rownames("protein")
    
    # calculate fold change to median of control condition
    wp_df <- (si_matrix / apply(si_matrix[, grepl(control_condition, colnames(si_matrix))], 1, median, na.rm =TRUE)) %>%
      rownames_to_column("protein") %>%
      # pivot to long format for plotting
      pivot_longer(cols = -protein,
                   names_to = "condition_replicate",
                   values_to = "channel_ratio") %>%
      # drop replicates where the protein was not quantified
      drop_na() %>%
      separate_wider_delim(
        cols =condition_replicate,
        delim = "_",
        names = c("condition", "technical_replicate", "experiment"),
        too_many = "drop"
      ) %>% mutate(
        residue = "C0",
        regulation = "Protein expression",
        reactivity_change = FALSE
      ) 
    
    si_matrix <- reactivity_table %>%
      filter(protein %in% proteins_to_plot) %>%
      # select columns needed for plot
      select(starts_with(c(
        conditions, c("protein", "residue", control_condition)
      ))) %>% mutate(residue = gsub(';', cysteine_seperator ,residue)) %>%
      # concat protein and residue
      unite(rowname, protein, residue) %>%
      column_to_rownames()
    # calculate fold change to median of control condition
    df <- (si_matrix / apply(si_matrix[, grepl(control_condition, colnames(si_matrix))], 1, median, na.rm =TRUE)) %>%
      rownames_to_column("protein_residue") %>%
      # pivot to long format for plotting
      pivot_longer(cols = -protein_residue,
                   names_to = "condition_replicate",
                   values_to = "channel_ratio") %>%
      # drop missing values
      drop_na() %>%
      separate_wider_delim(
        cols = c("protein_residue"),
        delim = "_",
        names = c("protein", "residue")
      ) %>%
      separate_wider_delim(
        cols = c("condition_replicate"),
        delim = "_",
        names = c("condition", "experiment", "technical_replicate"),
        too_many = "drop"
      ) %>% merge(
      y = full_df %>%
        select(protein, residue, reactivity_change, condition) %>%
        mutate(residue =gsub(",", cysteine_seperator, residue)),
      by = c("protein", "residue", "condition"),
      all.x = TRUE
    ) %>% mutate(regulation = "Reactivity") %>%
      rbind(wp_df) %>%
      mutate(
        # converting strings to factors so proteins and conditions have correct order
        protein = factor(protein, levels = proteins_to_plot),
        condition = factor(condition, levels = c(control_condition, conditions)),
        # get numeric index of cysteine for sorting
        residue_loc = str_split_fixed(residue, "C", 2)[, 2],
        residue_loc = str_split_fixed(residue_loc, cysteine_seperator, 2)[, 1] %>% as.numeric(),
        # format reactivity change boolean to string
        reactivity_change = as.character(reactivity_change), 
        reactivity_change = replace_na(reactivity_change, "FALSE")
      ) %>%
      filter(condition != "TLR9")
    
    # set plot dimensions
    # adding extra buffer on the bottom for annotation
    max_plot_value <- max(log2(df$channel_ratio)) * 1.1
    min_plot_value <- min(log2(df$channel_ratio)) * 1.2
    margin_value <- min_plot_value  + 0.05 * (max_plot_value - min_plot_value)
    
    plot <- df %>%
      unique() %>%
      ggplot(aes(
        x = reorder(residue, residue_loc),
        fill = condition,
        y = log2(channel_ratio),
        shape = regulation
      )) +
      geom_boxplot(
        outliers = FALSE,
        width = 0.7,
        lwd = LINE_WIDTH,
        show.legend = FALSE,
        alpha = 0.6,
        position = position_dodge(width = 0.8)
      ) +
      # replicate level LFC values
      geom_beeswarm(
        col = "black",
        stroke = POINT_STROKE,
        show.legend = FALSE,
        size = 0.9,
        alpha = 0.7,
        dodge.width = 0.8
      ) +
      facet_rep_grid(~ protein, scales = "free_x", space = "free", ) +
      my_theme() +
      theme(
        plot.title = element_text(
          size = 6,
          color = "black",
          family = "Helvetica",
          hjust = 0.5
        ),
        strip.background = element_blank(),
        panel.spacing = unit(2, "mm"),
        strip.text = element_text(
          size = 8,
          family = "Helvetica",
          color = "black"
        ),
        strip.clip = "off",
        panel.grid.major.y = element_line(linewidth = LINE_WIDTH, linetype = "solid", color = "grey90")
      ) +
      # boxes indicating reactivity changes
      geom_point(
        #data = rc_changes,
        mapping = aes(
          y = (min_plot_value + margin_value) / 2,
          color = condition,
          alpha = reactivity_change
        ),
        size = 1.5,
        show.legend = FALSE,
        shape = 22,
        position = position_dodge(width = 0.8),
        stroke = POINT_STROKE *1.5
      ) +
      # lines separating residues
      geom_vline(
        xintercept = 1.5:length(unique(df$residue)),
        size = LINE_WIDTH,
        linetype = "solid",
        color = "grey90"
      ) + 
      # margin line
      geom_hline(
        yintercept = margin_value,
        size = LINE_WIDTH,
        linetype = "solid",
        color = "black"
        )  +
      labs(y = substitute(paste('log'[2],"(FC from ", m, ")"),list(m = control_condition)),
           x = "",
           fill = "Reactivity") +
      scale_fill_manual(values = cols) +
      scale_alpha_discrete(range = c(0,1)) +
      scale_color_manual(values = cols) +
      scale_shape_manual(values = c("Reactivity" = 21,"Protein expression" = 24)) +
      scale_y_continuous(limits = c(min_plot_value, max_plot_value), expand = expansion(mult = c(0, 0)))
    
    print(plot)
    
    save_plot(
      paste0(results_dir, "/05_boxplots/", paste(proteins_to_plot, collapse = "_")),
      width = wd,
      height = ht
    )
    
    return(df)
  }
```


```{r}
# --- Helper functions ---

# Selects columns, filters proteins, computes fold change vs. control median,
# then pivots to long format and drops NAs
compute_fold_change <- function(mat, control_condition) {
  control_cols <- grepl(control_condition, colnames(mat))
  mat / apply(mat[, control_cols], 1, median, na.rm = TRUE)
}

prepare_matrix <- function(tbl, id_cols, conditions, control_condition, proteins_to_plot) {
  tbl %>%
    filter(protein %in% proteins_to_plot) %>%
    select(starts_with(c(conditions, id_cols, control_condition))) %>%
    unite(rowname, all_of(setdiff(id_cols, "protein")), remove = FALSE) %>%
    column_to_rownames("rowname")
}

pivot_to_long <- function(mat, value_col, control_condition) {
  compute_fold_change(mat, control_condition) %>%
    rownames_to_column("rowname") %>%
    pivot_longer(-rowname, names_to = "condition_replicate", values_to = value_col) %>%
    drop_na()
}

prep_combined_data <- function(whole_proteome_table, reactivity_table, proteins_to_plot, control_condition, conditions){
  cysteine_seperator <- ","
  # --- Whole-proteome (protein expression) data ---
  print("wp")
  wp_df <- whole_proteome_table %>%
    read_csv(show_col_types = FALSE) |> 
    filter(protein %in% proteins_to_plot) %>%
    select(starts_with(c(conditions, "protein", control_condition))) %>%
    column_to_rownames("protein") %>%
    pivot_to_long("channel_ratio", control_condition) %>%
    rename(protein = rowname) %>%
    separate_wider_delim(
      condition_replicate, "_",
      names = c("condition", "technical_replicate", "experiment"),
      too_many = "drop"
    ) %>%
    mutate(residue = "C0", regulation = "Protein expression", reactivity_change = FALSE)
  
  # --- Reactivity data ---
  print("rc")
  reactivity_df <- reactivity_table %>%
    read_csv(show_col_types = FALSE) |> 
    filter(protein %in% proteins_to_plot) %>%
    select(starts_with(c(conditions, "protein", "residue", control_condition))) %>%
    mutate(residue = gsub(";", cysteine_seperator, residue)) %>%
    unite(rowname, protein, residue) %>%
    column_to_rownames() %>%
    pivot_to_long("channel_ratio", control_condition) %>%
    separate_wider_delim(rowname, "_", names = c("protein", "residue")) %>%
    separate_wider_delim(
      condition_replicate, "_",
      names = c("condition", "experiment", "technical_replicate"),
      too_many = "drop"
    ) %>%
    left_join(
      full_df %>%
        select(protein, residue, reactivity_change, condition) %>%
        mutate(residue = gsub(", ", cysteine_seperator, residue)),
      by = c("protein", "residue", "condition")
    ) %>%
    mutate(regulation = "Reactivity")
  
  # --- Combine and finalise ---
  
  df <- bind_rows(reactivity_df, wp_df) %>%
    filter(condition != "TLR9") %>%
    mutate(
      protein          = factor(protein, levels = proteins_to_plot),
      condition        = factor(condition, levels = c(control_condition, conditions)),
      residue_loc      = str_split_fixed(residue, "C", 2)[, 2] %>%
                           str_split_fixed(cysteine_seperator, 2) %>% `[`(, 1) %>% as.numeric(),
      reactivity_change = replace_na(as.character(reactivity_change), "FALSE")
    )
  return(df)
}

# macrophage
macrophage_dir <- "/Users/henrysanford/dev/test_data/macrophage/"
macrophage_wp_table <- file.path(macrophage_dir, '01_whole_proteome/03_results/3_reps/03_combined_files/1/07_combfiles_percctrl_wp_min_2_replicates.csv')
macrophage_rc_table <- file.path(macrophage_dir, '02_reactivity/03_rc_analysis/20250225_3reps/04_combined_files/2/07_combfiles_cysaggr_percctrl_rc_min_2_replicates.csv')

# t cell
t_cell_dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Exhaustion manuscript/02_Mass-spectrometry data/'
t_cell_wp_table <- file.path(t_cell_dir, "01_whole_proteome/02_wp_analysis/20240821_6rep_2MC_two_replicate_var_filter/03_combined_files/20240821_6rep_2MC_two_replicate_var_filter/07_combfiles_percctrl_passes_two_replicate_variability_filter.csv")
t_cell_rc_table <- file.path(t_cell_dir, "02_reactivity/03_rc_analysis/20240821_5_rep_variability_filter/04_combined_files/20240821_5_rep_variability_filter/07_combfiles_cysaggr_percctrl_passes_two_replicate_variability_filter.csv")

protein_to_plot <- "GSR"

macrophage_df <- prep_combined_data(
  macrophage_wp_table,
  macrophage_rc_table,
  proteins_to_plot = protein_to_plot,
  control_condition = "M0",
  conditions = c("TLR4")
) |> mutate(dataset = "Macrophage")

t_cell_df <- prep_combined_data(
  t_cell_wp_table,
  t_cell_rc_table,
  proteins_to_plot = protein_to_plot,
  control_condition = "D2",
  conditions = c("D8C")
) |> mutate(dataset = "T cell")


```



```{r, fig.width=1.5, fig.height=2.3}
df <- macrophage_df |> 
  rbind(t_cell_df)

combined_color_palette <- c(cols, c("D2" = "#A1A1A1", "D4A" = "#ABA2D6", "D8A" = "#603EA6", "D4C" = "#FBB31B", "D8C" = "#E27753"))

# set plot dimensions
  # adding extra buffer on the bottom for annotation
  max_plot_value <- max(log2(df$channel_ratio)) * 1.1
  min_plot_value <- min(log2(df$channel_ratio)) * 1.1
  #margin_value <- min_plot_value  + 0.05 * (max_plot_value - min_plot_value)

df %>%
  mutate(residue = gsub("C0", "Expression", residue)) |> 
      unique() %>%
      ggplot(aes(
        x = reorder(residue, residue_loc),
        fill = condition,
        y = log2(channel_ratio),
        shape = regulation
      )) +
      geom_boxplot(
        outliers = FALSE,
        width = 0.7,
        lwd = LINE_WIDTH,
        show.legend = FALSE,
        alpha = 0.6,
        position = position_dodge(width = 0.8)
      ) +
      # replicate level LFC values
      geom_beeswarm(
        col = "black",
        stroke = POINT_STROKE,
        show.legend = FALSE,
        size = 0.9,
        alpha = 0.7,
        dodge.width = 0.8
      ) +
      facet_rep_grid(~ protein, scales = "free_x", space = "free", ) +
      my_theme() +
      theme(
        plot.title = element_text(
          size = 6,
          color = "black",
          family = "Helvetica",
          hjust = 0.5
        ),
        strip.background = element_blank(),
        panel.spacing = unit(2, "mm"),
        strip.text = element_text(
          size = 8,
          family = "Helvetica",
          color = "black"
        ),
        strip.clip = "off",
        panel.grid.major.y = element_line(linewidth = LINE_WIDTH, linetype = "solid", color = "grey90")
      ) +
      # lines separating residues
      geom_vline(
        xintercept = 1.5:length(unique(df$residue)),
        size = LINE_WIDTH,
        linetype = "solid",
        color = "grey90"
      ) + 
      labs(y = substitute(paste('log'[2],"(FC from ", m, ")"),list(m = "control")),
           x = "",
           fill = "Reactivity") +
  scale_fill_manual(values = combined_color_palette) +
      scale_alpha_discrete(range = c(0,1)) +
      scale_shape_manual(values = c("Reactivity" = 21,"Protein expression" = 24)) +
      scale_y_continuous(limits = c(min_plot_value, max_plot_value), expand = expansion(mult = c(0, 0))) + 
  facet_grid(cols = vars(dataset), space = "free", scales = "free_x") + 
        theme(axis.text.x = element_text(
          color = "black",
          angle = 45,
          size = FONT_SIZE,
          family = "Arial",
          vjust = 1,
          hjust = 1
        ))
save_plot(
      paste0(results_dir, "/05_boxplots/", "t_cell_mac_gsr"),
      width = 1.5,
      height = 2.3
    )

results_dir
results_dir
```





